---
title: "SMR SEER"
author: "Genesis Rodriguez"
date: "2024-12-20"
output: html_document
---

```{r libraries, include = FALSE}

library(readxl)
library(dplyr)
library(mgcv)
library(ggplot2)
library(openxlsx)
library(popEpi)
library(Epi)
library(data.table)  
library(lubridate)
library(stringr)

```

```{r cancer_data}

# Cleansed cancer data

cancer <- suppressWarnings(tryCatch({
  read.csv("C:/Users/Genesis Rodriguez/Box/Sanchez-Diaz Lab/CVD and BC/Registro Central de Cancer/Data/cancer_cleansed.csv")
}, error = function(e) {
  read.csv("~/Downloads/Registro Central de Cancer/cancer_cleansed.csv")
}, error = function(e) {
  read.csv("~/../../projects_shared/BBC/Project_Dra. Carola/Data/Registro Central de Cáncer/cancer_cleansed.csv")
}))
cancer <- cancer %>%
  mutate(FollowUpStart = as.Date(FollowUpStart),
         DiagnosisDate = as.Date(DiagnosisDate),
         FollowUpEnd = as.Date(FollowUpEnd),
         FollowUpEnd = if_else(month(FollowUpEnd) == 1 & day(FollowUpEnd) == 1,
                        FollowUpEnd + days(1),  # If the date is January 1, add one day
                        FollowUpEnd),
         maritallabel = str_to_title(ifelse(maritallabel == "unknown",
                               NA,
                               maritallabel)),
         Comorb.cat = factor(ifelse(Comorb == '0' , 'No', 
                             ifelse(as.numeric(Comorb) >= 1, 'Yes', 'Unknown')),
                             levels= c('Yes','No','Unknown')))

cancer$yearDiagCat <- cut(cancer$ydiag, 
                   breaks = c(2003, 2009, 2014, 2019), 
                   labels = c("2004-2009", "2010-2014", "2015-2019"),
                   right = TRUE)
cancer$deathYearCat <- cut(cancer$yy_dlc, 
                   breaks = c(2003, 2009, 2015, 2021, 2024), 
                   labels = c("2004-2009", "2010-2015", "2016-2021", "2022-2024"),
                   right = TRUE)

cancer_observed <- cancer %>%
  filter(FollowUpYears > 0)

```


```{r numerator}

cat("Global SMR")


# This function splits case-level observations by calendar time (per), age, and follow-up time
c <- lexpand(cancer_observed, status = cvdStatus != 0, 
             birth = birth_date,
             entry = FollowUpStart, # Without those that did not meet follow-up
             exit = FollowUpEnd, 
             breaks = list(per = 2004:2022, age = c(18, 60, 70, 80, Inf)), 
             aggre = list(agegroup = age, deathyear = per, ydiag, yearDiagCat, deathYearCat, Comorb.cat, maritallabel))

# Global SMR
smr_global <- sir(coh.data = c, coh.obs = 'from0to1', 
                  coh.pyrs = 'pyrs', 
                  ref.data = gen_2, 
                  ref.rate = 'gen_cvd_rate', 
                  adjust = c('deathyear', 'agegroup'), 
                  test.type = "homogeneity",
                  conf.type = "wald",
                  conf.level = 0.95); smr_global

