---
title: "SMR Calculations - US Reference Population"
date: "`r format(Sys.time(), '%b %d, %Y - %I:%M %p')`"
output:
  word_document: 
    reference_docx: "template.docx"
    fig_width: 9
    fig_height: 7
    fig_caption: true
editor_options:
  chunk_output_type: console
---

```{r libraries, include = FALSE, cache = FALSE}

knitr::opts_chunk$set(echo = FALSE, cache = FALSE, warning = FALSE, message = FALSE, results = "asis")

library(readxl)
library(dplyr)
library(mgcv)
library(ggplot2)
library(openxlsx)
library(popEpi) # SMR analysis
library(Epi) 
library(data.table)  
library(lubridate)
library(beepr) # sounds for R
library(stringr)
library(tidyr)
library(gtsummary)
library(gt) # for creating gt tables
library(webshot) # for saving images as png
library(survival) # for cumulative mortality km curves
library(forcats)
library(scales) # commas for numbers
library(survminer) #for ggsurvplot graphs
library(magick) # for tif file management
library(forestploter) # for forest plots
library(gridExtra) # joining forest plots


```

```{r functions}

# To suppress values under 10 for wide data frames

suppress_values <- function(data) {
  data %>%
    mutate(
      Observed = ifelse(as.numeric(Observed) < 10, "<10", comma(as.numeric(Observed))),
      Expected = ifelse(Observed == "<10", "-", comma(as.numeric(Expected), accuracy = 0.1)),
      SMR = ifelse(Observed == "<10", "-", sprintf("%.1f", as.numeric(SMR))),
      CI.Lower = ifelse(Observed == "<10", "-", sprintf("%.1f", as.numeric(CI.Lower))),
      CI.Upper = ifelse(Observed == "<10", "-", sprintf("%.1f", as.numeric(CI.Upper)))
    )
}

# Changing label of other race/ethnicity

recode_race_eth <- function(df) {
  df <- df %>%
    mutate(race_eth = case_when(
      race_eth == "Asian American, Native Hawaiian and other Pacific Islanders" ~ "AANHPI & AIAN",
      TRUE ~ as.character(race_eth)
    )) %>%
    mutate(race_eth = factor(race_eth, levels = c("PR Hispanic", "US Hispanic",
                                                "White", "Black", "AANHPI & AIAN")))
  return(df)
}

name_columns <- function(df) {
  race_ethnicity_pattern <- "_(White|Black|AANHPI & AIAN|US Hispanic|PR Hispanic)"
  
  setNames(
    str_replace_all(
      str_remove(colnames(df), race_ethnicity_pattern),
      c("Observed" = "O", "Expected" = "E")
    ),
    colnames(df)
  )
}


create_gt_table <- function(df, groupname_col = NULL) {
  if (!is.null(groupname_col)) {
    table <- df |> 
      gt(groupname_col = groupname_col)
  } else {
    table <- df |> 
      gt()
  }
  
  table <- table |> 
    spanners_and_header() |> 
    cols_label(.list = desired_colnames) |> 
    tab_style(
      style = cell_text(weight = "bold"),
      locations = cells_row_groups()) |> 
    cols_align(align = "center", columns = everything()) |> 
    tab_options(table.font.size = px(6)) |>
    tab_source_note(
      source_note = "SMRs were calculated as the observed number of cardiovascular disease deaths among breast cancer patients divided by the expected number of cardiovascular disease deaths in the race- and ethnic-matched general female U.S. population. SMR calculations for results with <10 events were suppressed. \n
      O = Observed;
      E = Expected; 
      SMR = Standarized Mortality Ratio; 
      AANHPI & AIAN encompasses: Asian American, Native Hawaiian and other Pacific Islanders along with American Indian/Alaska Natives."
    )
  
  return(table)
}


# Creating column names

spanners_and_header <- function(gt_tbl) {
  gt_tbl |> 
    tab_spanner(label = md('**White**'), columns = contains('White')) |> 
    tab_spanner(label = md('**Black**'), columns = contains('Black')) |> 
    tab_spanner(label = md('**AANHPI & AIAN**'), columns = contains('AANHPI & AIAN')) |> 
    tab_spanner(label = md('**US Hispanic**'), columns = contains('US Hispanic')) |> 
    tab_spanner(label = md('**PR Hispanic**'), columns = contains('PR Hispanic'))
}

# For quality control section: sums observations for each SEER*Stat table

sum_observed_columns <- function(df) {
  df %>%
    mutate(across(starts_with("Observed"), ~ as.numeric(gsub("<10", NA, .)))) %>%
    summarize(across(starts_with("Observed"), sum, na.rm = TRUE))
}

```

```{r cancer_data}

# Cleansed cancer data

cancer <- suppressWarnings(tryCatch({
  read.csv("C:/Users/Genesis Rodriguez/Box/Sanchez-Diaz Lab/CVD and BC/Registro Central de Cancer/Data/cancer_cleansed.csv")
}, error = function(e) {
  read.csv("~/Downloads/Registro Central de Cancer/cancer_cleansed.csv")
}, error = function(e) {
  read.csv("C:/Users/lcastro/OneDrive - Centro Comprensivo de Cancer UPR/projects_shared/BBC/Project_Dra. Carola/Data/Registro Central de Cáncer/cancer_cleansed.csv")
}))
cancer <- cancer %>%
  mutate(FollowUpStart = as.Date(FollowUpStart),
         DiagnosisDate = as.Date(DiagnosisDate),
         FollowUpEnd = as.Date(FollowUpEnd),
         maritallabel = str_to_title(ifelse(maritallabel == "unknown",
                               NA,
                               maritallabel)),
         Comorb.cat = factor(ifelse(Comorb == '0' , 'No', 
                             ifelse(as.numeric(Comorb) >= 1, 'Yes', 'Unknown')),
                             levels= c('Yes','No','Unknown')),
    Procedure = factor(Procedure,
                               levels = c("Surgery Only", "Chemotherapy Only",
                                          "Radiotherapy Only", "Chemotherapy and Radiotherapy",
                                          "Other")),
    Comorbidity = case_when(
    Comorb == -1 ~ NA_character_,
    Comorb == 0 ~ "0",
    Comorb == 1 ~ "1",
    Comorb == 2 ~ "2",
    Comorb >= 3 ~ "3+"
  ))

cancer_observed <- cancer %>%
  filter(FollowUpYears > 0)

cancer_observed$HealthRegion <- str_to_title(cancer_observed$HealthRegion)
cancer_observed$HealthRegion <- factor(cancer_observed$HealthRegion,
                            levels = c('Noreste Metro','Central Bayamon','Oeste Aguadilla/Mayaguez',
                                       'Sureste Caguas','Sur Ponce',"Norte Arecibo",
                                       'Este Fajardo','County Unknown'
                                       ))

```

```{r us_reference_race_pop}

# This version assumes that SEER*Stat is using the following file for referent rate classification:
# Rate File Name: U.S. Mortality 1975-2021 (Nov 2023 sub), Race (W,B,O), Event: COD rec (HIV grouped w/oth infectious)
# Created from NCHS Database: Mortality - All COD, Aggregated Total U.S. (1969-2022) <Katrina/Rita Population Adjustment>
# Adjusted by age, sex, year and race

# us_reference <- read.delim("~/Downloads/Statistical Analysis/Inputs/us_reference_by_race_wbo.txt") %>%
#   filter(!Year.of.death %in% "1969-2022",
#          Age.recode.with..1.year.olds != "Unknown",
#          Race.recode..White..Black..Other. != "Other unspecified (1978-1991)",
#          Year.of.death %in% 2004:2021) %>%
#   mutate(across(c(Age.Adjusted.Rate, Year.of.death, Count, Population), ~ as.numeric(gsub("[^0-9.]", "", .))),
#          agegroup = case_when(
#            str_detect(Age.recode.with..1.year.olds, "^(00|01|05|10|15|20|25|30|35)") ~ 0,
#            str_detect(Age.recode.with..1.year.olds, "^(40|45)") ~ 40,
#            str_detect(Age.recode.with..1.year.olds, "^(50|55)") ~ 50,
#            str_detect(Age.recode.with..1.year.olds, "^(60|65)") ~ 60,
#            str_detect(Age.recode.with..1.year.olds, "^(70|75)") ~ 70,
#            str_detect(Age.recode.with..1.year.olds, "^(80|85)") ~ 80,
#            TRUE ~ NA)) %>%
#   group_by(Year.of.death, agegroup, Race.recode..White..Black..Other.) %>%
#   summarise(n = sum(Count, na.rm = TRUE), Population = sum(Population, na.rm = TRUE), .groups = "drop") %>%
#   mutate(gen_cvd_rate = n / Population) %>%
#   filter(Race.recode..White..Black..Other. == "All races") %>%
#   rename(deathyear = Year.of.death, Race = Race.recode..White..Black..Other.)


```

```{r us_reference_race_eth}

# This version of us_reference comes from SEER*Stat's SMR files stratifies by age, sex, year and race.
# I reverse engineered the general population rates by dividing the expected number of deaths by the 
# person years for each level

us_reference_re <- read.csv("~/Downloads/Statistical Analysis/Inputs/smr_agegroup_deathyear_race_eth.txt") %>%
  select(Attained.Calendar.Year, Attained.Age, Race.Ethnicity.CVD, Observed, Expected, O.E, Person.Years.at.Risk) %>%
  filter(!Attained.Calendar.Year %in% c("2000-2004", "Total"), Attained.Age != "Total") %>%
  mutate(across(c(Observed, Expected, Person.Years.at.Risk), ~ as.numeric(gsub(",", "", .))),
         O.E = as.numeric(gsub("#", "", O.E)),
         agegroup = ifelse(Attained.Age != "0-39", as.numeric(substr(Attained.Age, 1, 2)), 0),
         deathyear = as.numeric(substr(Attained.Calendar.Year, 1, 4)),
         gen_cvd_rate = Expected / Person.Years.at.Risk)

us_reference_hispanics <- us_reference_re %>%
  filter(Race.Ethnicity.CVD == "US Hispanic")
```

```{r us_reference_pop}

us_reference_raw <- suppressWarnings(tryCatch({
  read.delim("C:/Users/Genesis Rodriguez/Downloads/Statistical Analysis/Inputs/us_reference.txt")
}, error = function(e) {
read.delim("~/Downloads/Statistical Analysis/Inputs/us_reference.txt")
  }, error = function(e) {
read.delim("C:/Users/lcastro/OneDrive - Centro Comprensivo de Cancer UPR/projects_shared/BBC/Project_Dra. Carola/Data/Inputs/us_reference.txt")
}))

us_reference_clean <- us_reference_raw %>%
  filter(!Year.of.death %in% c("1969-2022"),
          Cause.of.death.recode %in% c(
            "Diseases of Heart", 
            "Cerebrovascular Diseases", 
            "Atherosclerosis", 
            "Aortic Aneurysm and Dissection", 
            "Other Diseases of Arteries, Arterioles, Capillaries"),
         Age.recode.with..1.year.olds != "Unknown") %>%  
  mutate(across(
    c(Age.Adjusted.Rate, Count, Population, Std.Pop), ~ as.numeric(gsub("[^0-9.]", "", .))),
         across(c(2, 4, 5, 6, 7), as.numeric),     
         across(c(1,3, 4), as.factor)) %>%
  filter(Year.of.death %in% 2004:2021)

us_reference <- us_reference_clean %>%
#  filter(Sex == "Female") %>%
  mutate(age_group5 = as.numeric(substr(Age.recode.with..1.year.olds, 1, 2)),
    Age_group = case_when(
      str_detect(Age.recode.with..1.year.olds, "^00|^01|^05|^10|^15|^20|^25|^30|^35") ~ 0,
      str_detect(Age.recode.with..1.year.olds, "^40|^45") ~ 40,
      str_detect(Age.recode.with..1.year.olds, "^50|^55") ~ 50,
      str_detect(Age.recode.with..1.year.olds, "^60|^65") ~ 60,
      str_detect(Age.recode.with..1.year.olds, "^70|^75") ~ 70,
      str_detect(Age.recode.with..1.year.olds, "^80|^85") ~ 80,
      TRUE ~ NA
    )
  ) %>%
  group_by(Year.of.death, Age_group) %>%
  summarise(n = sum(Count),
            Population = sum(Population))

  # This comes from database:
# Mortality - All COD, Aggregated Total U.S. (1990-2022) <Katrina/Rita Population Adjustment>
# Adjusted by age, sex and year
us_reference <- read.delim("~/Downloads/Statistical Analysis/Inputs/us_reference.txt") %>%
  filter(!Year.of.death %in% "1990-2022",
         Cause.of.death.recode %in% c("Diseases of Heart", "Cerebrovascular Diseases", 
                                      "Atherosclerosis", "Aortic Aneurysm and Dissection", 
                                      "Other Diseases of Arteries, Arterioles, Capillaries"), # our definition of CVD
         Age.recode.with..1.year.olds != "Unknown",
         Year.of.death %in% 2004:2021) %>% # study period
  mutate(across(c(Age.Adjusted.Rate, Year.of.death, Count, Population), ~ as.numeric(gsub("[^0-9.]", "", .))),
         agegroup = case_when(
           str_detect(Age.recode.with..1.year.olds, "^(00|01|05|10|15|20|25|30|35)") ~ 0,
           str_detect(Age.recode.with..1.year.olds, "^(40|45)") ~ 40,
           str_detect(Age.recode.with..1.year.olds, "^(50|55)") ~ 50,
           str_detect(Age.recode.with..1.year.olds, "^(60|65)") ~ 60,
           str_detect(Age.recode.with..1.year.olds, "^(70|75)") ~ 70,
           str_detect(Age.recode.with..1.year.olds, "^(80|85)") ~ 80,
           TRUE ~ NA)) %>% # our population is 18+, but age groups are 0-39, 40-49, 50-59, 60-69, 70-79, 80+
  group_by(Year.of.death, agegroup) %>%
  summarise(n = sum(Count, na.rm = TRUE), Population = sum(Population, na.rm = TRUE), .groups = "drop") %>%
  mutate(gen_cvd_rate = n / Population) %>%
  rename(deathyear = Year.of.death)


```

```{r pr_description, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Descriptive Analysis ----------------------------------------------------

fisher_sim_p <- function(data, variable, by, ...) {
  result <- list()
  result$p <- stats::fisher.test(x = data %>% pull({{variable}}), y = data %>% pull({{by}}), simulate.p.value = T)$p.value
  result$test <- "Fisher's test with simulated p-value"
  result
}

gt_pr_table1 <- cancer_observed %>%
  select(c(Vital, AgeDx, AgeDxCat, FollowUpYears, AgeDeathCat2, SurvivalTime, HealthRegion, maritallabel, Procedure, StageAtDx)) %>%
  filter(Vital != "Alive", !is.na(AgeDxCat)) %>% 
 mutate(maritallabel = forcats::fct_na_value_to_level(maritallabel, level = "Unknown")) %>%
 tbl_summary(by = Vital,
              label = c(AgeDx ~ "Age at Diagnosis (Continuous)",
                        AgeDxCat ~ "Age at Diagnosis (Categorical)",
                        AgeDeathCat2 ~ "Age at Death",
                        SurvivalTime ~ "Time to death (years)",
                        HealthRegion ~ "Health Region",
                        maritallabel ~ "Marital Status",
                        Procedure ~ "Treatment",
                        StageAtDx ~ "Stage at Diagnosis",
                        AgeDxCat ~ "Age at Diagnosis (Categorical)"),
             type = all_continuous() ~ "continuous2",
             statistic = list(all_continuous() ~ c(
      "{mean}(±{sd})",
      "{median} ({p25}-{p75})",
      "{min}-{max}"),
    all_categorical() ~ "{n} ({p}%)"),
    digits = all_continuous() ~ 1) %>%
  add_overall(col_label = "**All death causes**  \nN = {style_number(N)}") %>%
add_p(
    test = list(
      all_continuous() ~ "oneway.test", # Welch's ANOVA due to no assumption of homogeneity of variances
      c(Procedure, HealthRegion) ~ "fisher_sim_p", # Simulated p-values
      all_categorical() ~ "chisq.test"  # Default chi-square for most variables
      
    )) %>%
  modify_footnote(everything() ~ NA) %>%
  modify_caption("**Table 1: Baseline Characteristics of Study Population**") %>%
  as_gt()

```

```{r pr_smrs}
c <- lexpand(cancer_observed, status = cvdStatus != 0, 
             birth = birth_date,
             entry = FollowUpStart, # Without those that did not meet follow-up
             exit = FollowUpEnd, 
             breaks = list(per = 2004:2022, age = c(0, 40, 50, 60, 70, 80, Inf)),
             aggre = list(agegroup = age, deathyear = per, ydiag, AgeDxCat, yearDiagCat, deathYearCat, Comorb.cat, Comorbidity, maritallabel, Procedure))

gen_2 <- suppressWarnings(tryCatch({
  read.csv("C:/Users/Genesis Rodriguez/Box/Sanchez-Diaz Lab/CVD and BC/Defunciones/cvm_rate_gen_pop2.csv")
}, error = function(e) {
  read.csv("~/Downloads/NCHS/cvm_rate_gen_pop2.csv")
}))
gen_2 <- as.data.table(gen_2)
setnames(gen_2, old = c("current_data_year", "AgeDeathCat2"), new = c("deathyear", "agegroup"))
gen_2 <- gen_2[, .(agegroup, deathyear, n, Population, gen_cvd_rate)]
gen_2[, agegroup := ifelse(agegroup == "80+", "80", 
                           ifelse(agegroup == "18-39", "0", 
                                sub("([0-9]+)-([0-9]+)", "\\1", agegroup)))]
gen_2[, agegroup := as.numeric(agegroup)]
gen_2[, gen_cvd_rate := gen_cvd_rate / 1000]

gt_pr_general <- left_join(gen_2, c, by = c("agegroup" = "agegroup", "deathyear" = "deathyear")) %>%
  mutate(expected = pyrs * gen_cvd_rate) %>%
  ungroup() %>%
  summarise(Observed = sum(from0to1, na.rm = TRUE), 
            Expected = round(sum(expected, na.rm = TRUE),1),
            pyrs = round(sum(pyrs, na.rm = TRUE),1),
            SMR = Observed/Expected,
            CI.Lower = qchisq(0.05 / 2, df = 2 * Observed) / (2 * Expected),
            CI.Upper = qchisq(1 - 0.05 / 2, df = 2 * (Observed + 1)) / (2 * Expected),
            SMR = paste0(sprintf("%.1f", SMR), "(", sprintf("%.1f", CI.Lower), "-", sprintf("%.1f", CI.Upper), "%)")) %>%
  rename("SMR (95%)" = SMR) %>%
  select(-c(CI.Lower, CI.Upper)) %>%
  gt()

# SMRs by year of death

# By year of death
smr_deathyear <- sir(coh.data = c, coh.obs = 'from0to1', 
                  coh.pyrs = 'pyrs', 
                  ref.data = gen_2, 
                  ref.rate = 'gen_cvd_rate', 
                  adjust = c('deathyear', 'agegroup'), 
                  test.type = "homogeneity",
                  conf.type = "wald",
                  conf.level = 0.95,
                  print ='deathyear') %>%
    mutate(sir.lo = qchisq(0.05 / 2, df = 2 * observed) / (2 * expected),
    sir.hi = qchisq(1 - 0.05 / 2, df = 2 * (observed + 1)) / (2 * expected))

f_prdeathyear <- ggplot(data = smr_deathyear[-1, ], aes(x = deathyear, y = sir)) +
  geom_line(color = "#4E79A7", size = 1.2) +  # Adjust line color and thickness
  geom_point(size = 3, shape = 21, fill = "white", color = "#4E79A7") +  # Add points with fill
  geom_ribbon(aes(ymin = sir.lo, ymax = sir.hi), fill = "lightblue", alpha = 0.3) +  # Confidence interval with fill color
  geom_hline(yintercept = 1, color = "red", linetype = "dashed", size = 1) +  # Red dashed line at SMR = 1
  geom_text(aes(label = round(sir, 1)), nudge_y = 0.7, nudge_x = -0.1, size = 4, color = "black") +  # Add labels with a slight upward offset
  xlab("Year of Death") +
  ylab("Standardized Mortality Ratio (SMR)") +
  scale_x_continuous(breaks = smr_deathyear$deathyear) +  # Add all year labels
  scale_y_continuous(breaks = seq(1, ceiling(max(smr_deathyear[-1, ]$sir.hi, na.rm = TRUE)), by = 2), limits = c(0, NA)) +
  theme_minimal(base_size = 14) + 
  theme(
    axis.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
    plot.title = element_text(face = "bold", hjust = 0.5))

# SMRs by grouped year of death

smr_deathYearCat <- smr_deathyear %>%
  mutate(`Year of Death` = case_when(
    deathyear >= 2005 & deathyear <= 2010 ~ "2005-2010",
    deathyear >= 2011 & deathyear <= 2016 ~ "2011-2016",
    deathyear >= 2017 & deathyear <= 2021 ~ "2017-2021"
  )) %>%
  group_by(`Year of Death`) %>%
  summarise(
    Observed = sum(observed),
    Expected = sum(expected),
    .groups = "drop"
  ) %>%
  mutate(SMR = Observed / Expected,
         CI.Lower = ifelse(
           Observed == 0, 0, 
           qchisq(0.05 / 2, df = 2 * Observed) / (2 * Expected)
         ),
         CI.Upper =  ifelse(
           Observed == 0, Inf, 
           qchisq(1 - 0.05 / 2, df = 2 * (Observed + 1)) / (2 * Expected))
         )

f_deathYearCat <- ggplot(data = smr_deathYearCat, aes(y = `Year of Death`, x = SMR, group = 1)) +
  geom_point(size = 4, color = "#4E79A7") +  # Blue points for SMR
  geom_errorbarh(aes(xmin = CI.Lower, xmax = CI.Upper), height = 0.2, color = "black") +  # Horizontal error bars
  geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 1) +  # Reference line at SMR = 1
  xlab("Standardized Mortality Ratio (SMR)") +
  geom_text(aes(label = round(SMR, 1)), nudge_y = 0.1, size = 4, color = "black") +
  scale_y_discrete(limits = rev(levels(smr_deathYearCat$`Year of Death`))) +  # Reverse y-axis order
  scale_x_continuous(
  breaks = scales::pretty_breaks(n = 6), 
  limits = c(.5, NA)) +
  ylab("Year of Death") +
  theme_minimal(base_size = 14) +  # Minimal theme
  theme(
    axis.title = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold"),  # Bold y-axis labels
    plot.title = element_text(face = "bold", hjust = 0.5))

# SMRs by year of diagnosis

# By year of diagnosis
smr_ydiag <- sir(coh.data = c, coh.obs = 'from0to1', 
                  coh.pyrs = 'pyrs', 
                  ref.data = gen_2, 
                  ref.rate = 'gen_cvd_rate', 
                  adjust = c('deathyear', 'agegroup'), 
                  test.type = "trend",
                  conf.type = "wald",
                  conf.level = 0.95,
                  print ='ydiag') %>%
    mutate(sir.lo = qchisq(0.05 / 2, df = 2 * observed) / (2 * expected),
    sir.hi = qchisq(1 - 0.05 / 2, df = 2 * (observed + 1)) / (2 * expected))

f_ydiag <- ggplot(data = smr_ydiag, aes(x = ydiag, y = sir)) +
  geom_line(color = "#4E79A7", size = 1.2) +  # Adjust line color and thickness
  geom_point(size = 3, shape = 21, fill = "white", color = "#4E79A7") +  # Add points with fill
  geom_ribbon(aes(ymin = sir.lo, ymax = sir.hi), fill = "lightblue", alpha = 0.3) +  # Confidence interval with fill color
  geom_text(aes(label = round(sir, 1)), nudge_y = 0.7, nudge_x = -0.1, size = 4, color = "black") +  # Add labels with a slight upward offset
  geom_hline(yintercept = 1, color = "red", linetype = "dashed", size = 1) +  # Red dashed line at SMR = 1
  xlab("Year of Diagnosis") +
  ylab("Standardized Mortality Ratio (SMR) (95% CI)") +
  scale_x_continuous(breaks = smr_ydiag$ydiag) +  # Add all year labels
  scale_y_continuous(breaks = seq(1, max(smr_ydiag$sir.hi, na.rm = TRUE), by = 2), limits = c(1, NA)) +  # Breaks start at 1
  theme_minimal(base_size = 14) +  # Use a minimal theme with larger font
  theme(
    axis.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
    plot.title = element_text(face = "bold", hjust = 0.5))

ggsave("/Users/genesisrodriguez/Downloads/Statistical Analysis/Outputs/Graphs/pr_diagnosisyear_graph.png", plot = f_ydiag, width = 10, height = 6, dpi = 300,  bg = "white")

f_ydiag2 <- ggplot(data = smr_ydiag, aes(x = sir, y = ydiag)) +
  geom_point(size = 4, color = "#4E79A7") +  # Blue points for SMR
  geom_errorbarh(aes(xmin = sir.lo, xmax = sir.hi), height = 0.2, color = "black") +  # Horizontal error bars
  geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 1) +  # Reference line at SMR = 1
  xlab("Standardized Mortality Ratio (SMR)") +
  scale_y_reverse(n.breaks = length(smr_ydiag$ydiag)) +  # Reverse the y-axis order for continuous variables
  scale_x_continuous(breaks = seq(1, max(smr_ydiag$sir.hi, na.rm = TRUE), by = 2), limits = c(1, NA)) +  # Breaks start at 1
  ylab("Year of Diagnosis") +
  theme_minimal(base_size = 14) +  # Minimal theme
  theme(
    axis.title = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold"),  # Bold y-axis labels
    plot.title = element_text(face = "bold", hjust = 0.5))

# SMRs by grouped year of diagnosis

smr_diagCat <- smr_ydiag %>%
  mutate(`Year of Diagnosis` = case_when(
    ydiag >= 2004 & ydiag <= 2009 ~ "2004-2009",
    ydiag >= 2010 & ydiag <= 2014 ~ "2010-2014",
    ydiag >= 2015 & ydiag <= 2019 ~ "2015-2019"
  )) %>%
  group_by(`Year of Diagnosis`) %>%
  summarise(
    Observed = sum(observed),
    Expected = sum(expected),
    .groups = "drop"
  ) %>%
  mutate(SMR = Observed / Expected,
         CI.Lower = ifelse(
           Observed == 0, 0, 
           qchisq(0.05 / 2, df = 2 * Observed) / (2 * Expected)
         ),
         CI.Upper =  ifelse(
           Observed == 0, Inf, 
           qchisq(1 - 0.05 / 2, df = 2 * (Observed + 1)) / (2 * Expected))
         )

f_ydiagcat <- ggplot(data = smr_diagCat, aes(x = SMR, y = `Year of Diagnosis`)) +
  geom_point(size = 4, color = "#4E79A7") +  # Blue points for SMR
  geom_errorbarh(aes(xmin = CI.Lower, xmax = CI.Upper), height = 0.2, color = "black") +  # Horizontal error bars
  geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 1) +  
  xlab("Standardized Mortality Ratio (SMR)") +
  ylab("Year of Diagnosis") +
  theme_minimal(base_size = 14) +  # Minimal theme
  theme(
    axis.title = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold"),  # Bold y-axis labels
    plot.title = element_text(face = "bold", hjust = 0.5)
  )

# SMRs by age group

# By age group
smr_agegroup <- sir(coh.data = c, coh.obs = 'from0to1', 
                  coh.pyrs = 'pyrs', 
                  ref.data = gen_2, 
                  ref.rate = 'gen_cvd_rate', 
                  adjust = c('deathyear', 'agegroup'), 
                  test.type = "homogeneity", # test for equal SMRs
                  conf.type = "wald",
                  conf.level = 0.95,
                  print ='AgeDxCat') %>%
    mutate(sir.lo = qchisq(0.05 / 2, df = 2 * observed) / (2 * expected),
    sir.hi = qchisq(1 - 0.05 / 2, df = 2 * (observed + 1)) / (2 * expected))

# smr_agegroup$agegroup <- factor(smr_agegroup$agegroup, 
#                                 levels = c(0, 40, 50, 60, 70, 80), 
#                                 labels = c("18-39", "40-49", "50-59", "60-69", "70-79", "80+"))

f_agegroup <- ggplot(data = smr_agegroup[smr_agegroup$AgeDxCat != "18-39", ], aes(x = AgeDxCat, y = sir, fill = AgeDxCat)) +
  geom_bar(stat = "identity", color = "black", width = 0.7) +  # Bar plot with black borders and custom width
  geom_errorbar(aes(ymin = sir.lo, ymax = sir.hi), width = 0.2, color = "black") +  # Add error bars
  geom_text(aes(label = sprintf("%.1f", sir)), vjust = -0.5, hjust = -0.6, size = 5) +  # Add SMR labels above bars
  xlab("Age at Diagnosis") +
  ylab("Standardized Mortality Ratio (SMR) (95% CI)") +
  scale_fill_manual(values = rep("indianred", length(unique(smr_agegroup$AgeDxCat)))) + 
  theme_minimal() +
  theme(text = element_text(size = 16),
    axis.title = element_text(face = "bold"),
    axis.text.x = element_text(hjust = 1),  # Rotate x-axis labels
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "none"  # Remove legend if not needed
  )

ggsave("/Users/genesisrodriguez/Downloads/Statistical Analysis/Outputs/Graphs/pr_agegroup_diagnosis_graph.png", plot = f_agegroup, width = 11, height = 5.5, dpi = 300,  bg = "white")

# By treatment

smr_treatment <- sir(coh.data = c, coh.obs = 'from0to1',
                coh.pyrs = 'pyrs',
                ref.data = gen_2,
                ref.rate = 'gen_cvd_rate',
                adjust = c('deathyear', 'agegroup'),
                test.type = "homogeneity",
                conf.type = "wald",
                conf.level = 0.95,
                print ='Procedure') %>%
    mutate(sir.lo = qchisq(0.05 / 2, df = 2 * observed) / (2 * expected),
    sir.hi = qchisq(1 - 0.05 / 2, df = 2 * (observed + 1)) / (2 * expected))

f_treatment <- ggplot(data = smr_treatment, aes(x = Procedure, y = sir, fill = Procedure)) +
  geom_bar(stat = "identity", color = "black", width = 0.7) +  # Bar plot with black borders and custom width
  geom_errorbar(aes(ymin = sir.lo, ymax = sir.hi), width = 0.2, color = "black") +  # Add error bars
  geom_text(aes(label = sprintf("%.1f", sir)), vjust = -0.5, hjust = -0.6, size = 5) +  # Add SMR labels above bars
  xlab("Type of Treatment") +
  ylab("Standardized Mortality Ratio (SMR) (95% CI)") +
  scale_fill_manual(values = rep("indianred", length(unique(smr_treatment$Procedure)))) + 
  theme_minimal() + 
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +  # Wrap x-axis labels to fit within ~10 characters per line
  theme(text = element_text(size = 16, angle = 0, hjust = 0.5),
    axis.title = element_text(face = "bold"),
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "none")

ggsave("/Users/genesisrodriguez/Downloads/Statistical Analysis/Outputs/Graphs/pr_treatment.tif", plot = f_treatment, width = 11, height = 6, dpi = 300,  bg = "white")

# By comorbidity status

smr_comorb <- sir(coh.data = c, coh.obs = 'from0to1',
                coh.pyrs = 'pyrs',
                ref.data = gen_2,
                ref.rate = 'gen_cvd_rate',
                adjust = c('deathyear', 'agegroup'),
                test.type = "homogeneity",
                conf.type = "wald",
                conf.level = 0.95,
                print ='Comorbidity') %>%
    mutate(sir.lo = qchisq(0.05 / 2, df = 2 * observed) / (2 * expected),
    sir.hi = qchisq(1 - 0.05 / 2, df = 2 * (observed + 1)) / (2 * expected))

f_comorb <- ggplot(data = smr_comorb, aes(x = sir, y = fct_rev(Comorbidity))) +
  geom_point(size = 4, color = "#4E79A7") +  # Blue points for SMR
  geom_errorbarh(aes(xmin = sir.lo, xmax = sir.hi), height = 0.2, color = "black") +  # Horizontal error bars
  geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 1) +  # Reference line at SMR = 1
  geom_text(aes(label = sprintf("%.1f", sir)), vjust = -0.7, size = 5) +  # Add SMR labels above bars
  xlab("Standardized Mortality Ratio (SMR)") +
  scale_y_discrete(name = "Number of comorbidities") +  # Discrete scale for y-axis
  scale_x_continuous(breaks = seq(1, max(smr_comorb$sir.hi, na.rm = TRUE), by = 2), limits = c(1, NA)) +  # Breaks start at 1
  theme_minimal(base_size = 14) +  # Minimal theme
  theme(
    axis.title = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold"),  # Bold y-axis labels
    plot.title = element_text(face = "bold", hjust = 0.5)
  )

ggsave("/Users/genesisrodriguez/Downloads/Statistical Analysis/Outputs/Graphs/pr_comorb.tif", plot = f_comorb, width = 11, height = 6, dpi = 300,  bg = "white")

```

```{r sir, include = FALSE}

# SIR session

pr_sir_general <- cancer_observed %>%
  mutate(Vital = as.character(Vital),
    `Cause of Death` = case_when(
      Vital == "CVD death" ~ "CVD Death",
      Vital == "BC death" ~ "Other Causes of Death",
      Vital == "Other Death" ~ "Other Causes of Death",
      TRUE ~ Vital)) %>%
  group_by(`Cause of Death`) %>%
  count(name = "Count") %>%
  mutate(race_eth = "PR Hispanic")

us_sir_general <- suppressWarnings(tryCatch({
  read.delim("~/Downloads/Statistical Analysis/Inputs/sir_general.txt")
}, error = function(e) {
read.delim("C:/Users/lcastro/OneDrive - Centro Comprensivo de Cancer UPR/projects_shared/BBC/Project_Dra. Carola/Data/Inputs/sir_general.txt")
  } )) %>%
select(Index.Record.Cause.of.Death.CVD.recode,
       Index.Record.Race.Ethnicity.CVD, 
       Count) %>%
  rename(`Cause of Death` = Index.Record.Cause.of.Death.CVD.recode,
         race_eth = Index.Record.Race.Ethnicity.CVD) %>% 
  relocate(race_eth) %>%
  mutate(Count = as.numeric(gsub(",", "", Count)),
        `Cause of Death` = gsub("\\[OTHER\\]", "Other Causes of Death",
    `Cause of Death`),
    `Cause of Death` = gsub("CVD", "CVD Death",
    `Cause of Death`))

us_sir_general <- recode_race_eth(us_sir_general)

us_sir_general_wider <- us_sir_general %>%
 pivot_wider(names_from = c(race_eth), values_from = Count)


merged_sir_general <- rbind(pr_sir_general, us_sir_general) %>%
  mutate(race_eth = factor(race_eth, levels = c("PR Hispanic", "US Hispanic",
                                                "White", "Black", "AANHPI & AIAN"))) %>%
  group_by(race_eth) %>%
  mutate(Count = paste0(comma(Count), " (", sprintf("%.1f", Count/ sum(Count) * 100), "%)")) %>%
  arrange(race_eth) 

merged_sir_general_wider <- merged_sir_general %>%
  pivot_wider(names_from = c(race_eth), values_from = Count)

# Creating gt table
  
gt_sir_general <- merged_sir_general_wider |> 
  rename(`Vital status` = `Cause of Death`) |>
  gt() |> 
    tab_style(
      style = cell_text(weight = "bold"), 
      locations = cells_column_labels(everything())) |> 
  cols_align(align = "center", columns = everything()); #gt_sir_general

# US sample size
us_n <- us_sir_general %>% summarise(sum = sum(Count))

```

```{r sir_case_listing, include = FALSE}

us_sir_general_cl_df <- read.delim2("~/Downloads/Statistical Analysis/Inputs/sir_general_case_listing.txt") %>%
  mutate(Survival.months = as.numeric(Survival.months),
         FollowUpYears = round((Survival.months - 12) / 12, 2),
         cvdStatus = ifelse(Cause.of.Death.CVD.recode == "CVD", 1, 0)) %>%
  filter(Event.Number == "Index Record",
         Race.Ethnicity.CVD != " ")

# US General SIR table
#us_sir_general %>% arrange(`Cause of Death`)
us_sir_general_cl <- us_sir_general_cl_df %>%
  count(Cause.of.Death.CVD.recode, Race.Ethnicity.CVD, name = "Count") %>%
  arrange(Cause.of.Death.CVD.recode)

# Total U.S. CVD Deaths
#us_sir_general %>% filter(`Cause of Death` == "CVD Death") %>% summarise(sum = sum(Count))
#us_sir_general_cl %>% filter(Cause.of.Death.CVD.recode == "CVD") %>% nrow()

# Total U.S. Persons
#us_sir_general %>% summarise(sum = sum(Count))
#us_sir_general_cl %>% nrow()

# U.S. case listing by vital status
#us_sir_general %>% group_by(`Cause of Death`) %>% summarise(n = sum(Count))
#table(us_sir_general_cl$Cause.of.Death.CVD.recode)

```

```{r}

source("~/BC_CVD/flowChart.R")

```

```{r global}

# PR SMR
c <- lexpand(cancer_observed, status = cvdStatus != 0, 
             birth = birth_date,
             entry = FollowUpStart,
             exit = FollowUpEnd, 
             breaks = list(per = 2004:2022, age = c(0, 40, 50, 60, 70, 80, Inf)),
             aggre = list(agegroup = age, deathyear = per))

pr_smr_general <- sir(coh.data = c, coh.obs = 'from0to1', 
                  coh.pyrs = 'pyrs', 
                  ref.data = us_reference, 
                  ref.rate = 'gen_cvd_rate', 
                  adjust = c('deathyear', 'agegroup'), 
                  test.type = "homogeneity",
                  conf.type = "univariate",
                  conf.level = 0.95) %>%
  as_tibble() %>%
  mutate(race_eth = "PR Hispanic") %>%
  select(-c(pyrs, p_value)) %>%
  rename(Observed = observed,
         Expected = expected,
         SMR = sir,
        CI.Lower = sir.lo,
        CI.Upper = sir.hi) %>%
   mutate(CI.Lower = qchisq(0.05 / 2, df = 2 * Observed) / (2 * Expected),
    CI.Upper = qchisq(1 - 0.05 / 2, df = 2 * (Observed + 1)) / (2 * Expected))  %>% 
  relocate(race_eth)

us_smr_general <- suppressWarnings(tryCatch({
  read.delim("~/Downloads/Statistical Analysis/Inputs/smr_general.txt")
}, error = function(e) {
read.delim("C:/Users/lcastro/OneDrive - Centro Comprensivo de Cancer UPR/projects_shared/BBC/Project_Dra. Carola/Data/Inputs/smr_general.txt")
})) %>%
select(Race.Ethnicity.CVD, 
         Observed, Expected, O.E, CI.Lower, CI.Upper) %>%
  rename(race_eth = Race.Ethnicity.CVD,
         SMR = O.E) %>% 
  relocate(race_eth) %>%
  mutate(SMR = gsub("#", "", SMR),
         Observed = gsub(",", "", Observed),
         Expected = gsub(",", "", Expected))

merged_general <- rbind(pr_smr_general, us_smr_general) %>%
  mutate(race_eth = as.factor(race_eth),
         across(c(Expected:CI.Upper), ~ format(round(as.numeric(.), 1), nsmall = 1)),
         race_eth = factor(race_eth, levels = c("PR Hispanic", "US Hispanic",
                                                "White", "Black", "Asian American, Native Hawaiian and other Pacific Islanders"))) %>%
  arrange(race_eth)

# Supressing values under 10

merged_general <- suppress_values(merged_general)

# Renaming other race/ethnicity

merged_general <- recode_race_eth(merged_general)

# Merging US and PR General
merged_general_wider <- merged_general %>%
  mutate("95% CI" = ifelse(Observed != "<10",
         paste0("(", CI.Lower, " - ", CI.Upper, ")"),
         "-")) %>%
  select(-c(CI.Lower, CI.Upper)) %>%
  pivot_wider(names_from = c(race_eth), values_from = c(Observed, Expected, SMR, "95% CI")) |> 
  arrange("Age Group")

# Setting column names
desired_colnames <- name_columns(merged_general_wider)

# Creating gt table
gt_general <- create_gt_table(merged_general_wider)

f_general <- ggplot(merged_general, aes(y = fct_rev(race_eth), x = as.numeric(SMR))) +
    geom_point(size = 2, color = "#4E79A7") +  # Plot SMR points
    geom_errorbar(aes(xmin = as.numeric(CI.Lower), xmax = as.numeric(CI.Upper)), width = 0.2) +
    theme_minimal() +
    geom_text(aes(label = round(as.numeric(SMR), 1)), position = position_dodge(0), vjust = -1) +
    labs(x = "Standardized Mortality Ratio (SMR)", y = "Race/Ethnicity") +
    theme(axis.text.x = element_text(hjust = 1)) +
    scale_y_discrete(labels = function(x) str_wrap(x, width = 20))


ggsave("/Users/genesisrodriguez/Downloads/Statistical Analysis/Outputs/Graphs/general_graph.tif", plot = f_general, width = 10, height = 8, dpi = 300,  bg = "white")

```

```{r deathyear_table}

# PR SMRs by Death Year
c <- lexpand(cancer_observed, status = cvdStatus != 0, 
             birth = birth_date,
             entry = FollowUpStart, # Without those that did not meet follow-up
             exit = FollowUpEnd, 
             breaks = list(per = 2004:2022, age = c(0, 40, 50, 60, 70, 80, Inf)),
             aggre = list(agegroup = age, deathyear = per))

pr_smr_deathyear <- sir(coh.data = c, coh.obs = 'from0to1', 
                  coh.pyrs = 'pyrs', 
                  ref.data = us_reference, 
                  ref.rate = 'gen_cvd_rate', 
                  adjust = c('deathyear', 'agegroup'), 
                  test.type = "homogeneity",
                  conf.type = "univariate",
                  conf.level = 0.95, print = "deathyear") %>%
  as_tibble() %>%
  mutate(race_eth = "PR Hispanic") %>%
  select(-c(pyrs, p_value)) %>%
  rename("Year of Death" = deathyear,
         Observed = observed,
         Expected = expected,
         SMR = sir,
        CI.Lower = sir.lo,
        CI.Upper = sir.hi) %>%
   mutate(CI.Lower = qchisq(0.05 / 2, df = 2 * Observed) / (2 * Expected),
    CI.Upper = qchisq(1 - 0.05 / 2, df = 2 * (Observed + 1)) / (2 * Expected))  %>% 
  relocate(race_eth)

# US SMRs by Death Year
us_smr_deathyear <- suppressWarnings(tryCatch({
  read.delim("~/Downloads/Statistical Analysis/Inputs/smr_deathyear.txt")
}, error = function(e) {
read.delim("C:/Users/lcastro/OneDrive - Centro Comprensivo de Cancer UPR/projects_shared/BBC/Project_Dra. Carola/Data/Inputs/smr_deathyear.txt")
})) %>%
select(c(Year.of.death.recode,
  Race.Ethnicity.CVD, 
         Observed, Expected, O.E, CI.Lower, CI.Upper)) %>%
  filter(!(Year.of.death.recode %in% c("2000-2021", "2000", "2001", "2002", "2003",
                                       "2004", "Alive at last contact")),
                  Race.Ethnicity.CVD != "All races/ethnicities") %>%
  rename("Year of Death" = Year.of.death.recode, 
         race_eth = Race.Ethnicity.CVD,
         SMR = O.E) %>% 
  relocate(race_eth) %>%
  mutate(SMR = gsub("#", "", SMR),
         Observed = gsub(",", "", Observed),
         Expected = gsub(",", "", Expected))

# Merging PR and US by Death Year 

merged_deathyear <- rbind(pr_smr_deathyear, us_smr_deathyear) %>%
  mutate(Observed = as.numeric(Observed),
    across(c(race_eth, "Year of Death"), as.factor),
         race_eth = factor(race_eth, levels = c("PR Hispanic", "US Hispanic",
                                                "White", "Black", "Asian American, Native Hawaiian and other Pacific Islanders"))) %>%
  arrange(race_eth)

# Suppressing values under 10

merged_deathyear <- suppress_values(merged_deathyear)

# Renaming other race/ethnicity

merged_deathyear <- recode_race_eth(merged_deathyear)

# Widening by Race/Ethnicity

merged_deathyear_wider <- merged_deathyear %>%
 mutate("95% CI" = ifelse(Observed != "<10",
         paste0("(", CI.Lower, " - ", CI.Upper, ")"),
         "-")) %>%
  select(-c(CI.Lower, CI.Upper)) %>%
  pivot_wider(names_from = c(race_eth), values_from = c(Observed, Expected, SMR, "95% CI")) |> 
  arrange("Year of Death")

# Naming columns
desired_colnames <- name_columns(merged_deathyear_wider)

# Creating gt table

gt_deathyear <- create_gt_table(merged_deathyear_wider)

# Saving table
gtsave(gt_deathyear, "~/Downloads/Statistical Analysis/Outputs/Tables/DeathYear_table.html")

# webshot::webshot("/Users/genesisrodriguez/Downloads/Statistical Analysis/Outputs/Tables/DeathYear_table.html",
#                  path.expand("~/Downloads/Statistical Analysis/Outputs/Tables/DeathYear_table.png"),
#                  vwidth = 2450,
#                  vheight = 1300)

```

```{r deathyear_graph}

# Create the ggplot
f_deathyear <- ggplot(merged_deathyear, aes(x = `Year of Death`, y = as.numeric(SMR), color = race_eth, fill = race_eth, group = race_eth)) +
  geom_line(size = 1) +  # Line for SMR over years
 # geom_ribbon(aes(ymin = as.numeric(CI.Lower), ymax = as.numeric(CI.Upper)), alpha = 0.2, color = NA) +  # Confidence interval ribbons
  theme_minimal() +
  geom_hline(yintercept = 1, linetype = "solid", color = "black", size = 0.5) +  # Horizontal line at SMR = 1
  labs(
    x = "Year of Death",
    y = "SMR (95% CI)",
    color = "Race/Ethnicity",
    fill = "Race/Ethnicity"
  ) +
  scale_y_continuous(limits = c(0, 35)) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  guides(
    color = guide_legend(nrow = 2),
    fill = guide_legend(nrow = 2)
  ) +
  theme(
    text = element_text(size = 10),
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    legend.position = "bottom",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9))

ggsave("/Users/genesisrodriguez/Downloads/Statistical Analysis/Outputs/Graphs/grouped_deathyear_graph.tif", plot = f_deathyear, width = 10, height = 8, dpi = 300,  bg = "white")

```

```{r grouped_deathyear_table}

 pr_smr_g_deathyear <- pr_smr_deathyear %>%
  mutate(`Year of Death` = case_when(
    `Year of Death` >= 2005 & `Year of Death` <= 2010 ~ "2005-2010",
    `Year of Death` >= 2011 & `Year of Death` <= 2016 ~ "2011-2016",
    `Year of Death` >= 2017 & `Year of Death` <= 2021 ~ "2017-2021"
  )) %>%
  group_by(race_eth, `Year of Death`) %>%
  summarise(
    Observed = sum(Observed),
    Expected = sum(Expected),
    .groups = "drop"
  ) %>%
  mutate(SMR = Observed / Expected,
         CI.Lower = ifelse(
           Observed == 0, 0, 
           qchisq(0.05 / 2, df = 2 * Observed) / (2 * Expected)
         ),
         CI.Upper =  ifelse(
           Observed == 0, Inf, 
           qchisq(1 - 0.05 / 2, df = 2 * (Observed + 1)) / (2 * Expected))
         )

# US SMRs by Death Year
us_smr_g_deathyear <- us_smr_deathyear %>%
  mutate(`Year of Death` = case_when(
    `Year of Death` >= 2005 & `Year of Death` <= 2010 ~ "2005-2010",
    `Year of Death` >= 2011 & `Year of Death` <= 2016 ~ "2011-2016",
    `Year of Death` >= 2017 & `Year of Death` <= 2021 ~ "2017-2021"
  ),
  Observed = as.numeric(Observed),
  Expected = as.numeric(Expected)) %>%
  group_by(race_eth, `Year of Death`) %>%
   summarise(
    Observed = sum(Observed),
    Expected = sum(Expected),
    .groups = "drop") %>%
  mutate(SMR = Observed / Expected,
         CI.Lower = ifelse(
           Observed == 0, 0, 
           SMR * exp(-1.96 * sqrt(1 / Observed))
         ),
         CI.Upper =  ifelse(
           Observed == 0, Inf, 
           SMR * exp(1.96 * sqrt(1 / Observed)))) %>%
  arrange(race_eth, `Year of Death`)

# Merging PR and US by Death Year 

merged_g_deathyear <- rbind(pr_smr_g_deathyear, us_smr_g_deathyear) %>%
  mutate(`Year of Death` = as.factor(`Year of Death`),
         race_eth = factor(race_eth, levels = c("PR Hispanic", "US Hispanic", "White", "Black", "Asian American, Native Hawaiian and other Pacific Islanders")),
         across(c(Expected:CI.Upper), ~ format(round(as.numeric(.), 1), nsmall = 1))) %>%
  arrange(race_eth)

# Suppressing values under 10

merged_g_deathyear <- suppress_values(merged_g_deathyear)


# Renaming other race/ethnicity

merged_g_deathyear <- recode_race_eth(merged_g_deathyear)

# Widening by Race/Ethnicity

merged_g_deathyear_wider <- merged_g_deathyear %>%
 mutate("95% CI" = ifelse(Observed != "<10",
         paste0("(", CI.Lower, " - ", CI.Upper, ")"),
         "-")) %>%
  select(-c(CI.Lower, CI.Upper)) %>%
  pivot_wider(names_from = c(race_eth), values_from = c(Observed, Expected, SMR, "95% CI")) |> 
  arrange("Year of Death")

# Naming columns
desired_colnames <- name_columns(merged_g_deathyear_wider)

# Creating gt table

gt_g_deathyear <- create_gt_table(merged_g_deathyear_wider)

# Saving table
gtsave(gt_deathyear, "~/Downloads/Statistical Analysis/Outputs/Tables/Grouped_DeathYear_table.html")

# webshot::webshot("/Users/genesisrodriguez/Downloads/Statistical Analysis/Outputs/Tables/Grouped_DeathYear_table.html",
#                  path.expand("~/Downloads/Statistical Analysis/Outputs/Tables/Grouped_DeathYear_table.png"),
#                  vwidth = 2450,
#                  vheight = 1300)

```

```{r grouped_deathyear_graph}

f_g_deathyear <- ggplot(merged_g_deathyear, aes(x = `Year of Death`, y = as.numeric(SMR))) +
    geom_point(size = 0.5, position = position_dodge(width = 0.5)) +
    geom_errorbar(
        aes(ymin = as.numeric(CI.Lower), ymax = as.numeric(CI.Upper)),
        width = 0.2, position = position_dodge(width = 0.5)) +
    geom_hline(yintercept = 1, linetype = "solid", color = "red", size = 0.2) +
    facet_wrap(~ race_eth, ncol = 2) +  # Adjust `ncol` to create square facets
    theme_bw() +
    labs(x = "", y = "SMR (95% CI)") +
    theme(
        text = element_text(size = 7), 
        strip.text = element_text(angle = 0, face = "bold", size = 6), 
        strip.background = element_rect(fill = NA, color = NA),
        axis.text.x = element_text(size = 6),  
        axis.text.y = element_text(size = 6), 
        legend.position = "none",
        panel.spacing = unit(0.5, "lines"))


ggsave("/Users/genesisrodriguez/Downloads/Statistical Analysis/Outputs/Graphs/g_deathyear.tif", plot = f_g_deathyear, width = 10, height = 8, dpi = 300,  bg = "white")

```

```{r diagnosis_table}

# PR SMRs by Diagnosis Year
c <- lexpand(cancer_observed, status = cvdStatus != 0, 
             birth = birth_date,
             entry = FollowUpStart, # Without those that did not meet follow-up
             exit = FollowUpEnd, 
             breaks = list(per = 2004:2022, age = c(0, 40, 50, 60, 70, 80, Inf)),
             aggre = list(agegroup = age, deathyear = per, ydiag))

pr_smr_diagnosisyear <- sir(coh.data = c, coh.obs = 'from0to1', 
                  coh.pyrs = 'pyrs', 
                  ref.data = us_reference, 
                  ref.rate = 'gen_cvd_rate', 
                  adjust = c('deathyear', 'agegroup'), 
                  test.type = "homogeneity",
                  conf.type = "univariate",
                  conf.level = 0.95, print = "ydiag") %>%
  as_tibble() %>%
  mutate(race_eth = "PR Hispanic") %>%
  select(-c(pyrs, p_value)) %>%
  rename("Year of Diagnosis" = ydiag,
         Observed = observed,
         Expected = expected,
         SMR = sir,
        CI.Lower = sir.lo,
        CI.Upper = sir.hi) %>%
  mutate(CI.Lower = qchisq(0.05 / 2, df = 2 * Observed) / (2 * Expected),
         CI.Upper = qchisq(1 - 0.05 / 2, df = 2 * (Observed + 1)) / (2 * Expected))  %>% 
  relocate(race_eth)

# US SMRs by Diagnosis Year
us_smr_diagnosisyear <- suppressWarnings(tryCatch({
  read.delim("~/Downloads/Statistical Analysis/Inputs/smr_diagnosisyear.txt")
}, error = function(e) {
read.delim("C:/Users/lcastro/OneDrive - Centro Comprensivo de Cancer UPR/projects_shared/BBC/Project_Dra. Carola/Data/Inputs/smr_diagnosisyear.txt")
})) %>% select(c(Year.of.diagnosis,
  Race.Ethnicity.CVD, 
         Observed, Expected, O.E, CI.Lower, CI.Upper)) %>%
  filter(!(Year.of.diagnosis %in% c("2000-2021", "2000", "2001", "2002", "2003", "2020", "2021")),
                  Race.Ethnicity.CVD != "All races/ethnicities") %>%
  rename("Year of Diagnosis" = Year.of.diagnosis, 
         race_eth = Race.Ethnicity.CVD,
         SMR = O.E) %>% 
  relocate(race_eth) %>%
  mutate(SMR = gsub("#", "", SMR),
         Observed = gsub(",", "", Observed),
         Expected = gsub(",", "", Expected))

# Merging PR and US by Death Year 

merged_diagnosisyear <- rbind(pr_smr_diagnosisyear, us_smr_diagnosisyear) %>%
  mutate(across(c(race_eth, "Year of Diagnosis"), as.factor),
         across(c(Expected:CI.Upper), ~ format(round(as.numeric(.), 1), nsmall = 1)),
         race_eth = factor(race_eth, levels = c("PR Hispanic", "US Hispanic",
                                                "White", "Black", "Asian American, Native Hawaiian and other Pacific Islanders"))) %>%
  arrange(race_eth)

# Suppressing values under 10

merged_diagnosisyear <- suppress_values(merged_diagnosisyear)


# Renaming other race/ethnicity

merged_diagnosisyear <- recode_race_eth(merged_diagnosisyear)

# Widening by Race/Ethnicity

merged_diagnosisyear_wider <- merged_diagnosisyear %>%
 mutate("95% CI" = ifelse(Observed != "<10",
         paste0("(", CI.Lower, " - ", CI.Upper, ")"),
         "-")) %>%
  select(-c(CI.Lower, CI.Upper)) %>%
  pivot_wider(names_from = c(race_eth), values_from = c(Observed, Expected, SMR, "95% CI")) |> 
  arrange("Year of Diagnosis")

# Naming columns
desired_colnames <- name_columns(merged_diagnosisyear_wider)

# Creating gt table

gt_diagnosisyear <- create_gt_table(merged_diagnosisyear_wider)

# Saving table
gtsave(gt_diagnosisyear, "~/Downloads/Statistical Analysis/Outputs/Tables/DiagnosisYear_table.html")

# webshot::webshot("/Users/genesisrodriguez/Downloads/Statistical Analysis/Outputs/Tables/DiagnosisYear_table.html",
#                  path.expand("~/Downloads/Statistical Analysis/Outputs/Tables/DiagnosisYear_table.png"),
#                  vwidth = 2450,
#                  vheight = 700)

```

```{r diagnosisyear_graph}

# Create the ggplot
f_diagnosisyear <- ggplot(merged_diagnosisyear, aes(x = `Year of Diagnosis`, y = as.numeric(SMR), color = race_eth, fill = race_eth, group = race_eth)) +
  geom_line(size = 1) +  # Line for SMR over years
  theme_minimal() +
  geom_hline(yintercept = 1, linetype = "solid", color = "black", size = 0.5) +  # Horizontal line at SMR = 1
  labs(
    x = "Year of Diagnosis",
    y = "SMR (95% CI)",
    color = "Race/Ethnicity",
    fill = "Race/Ethnicity") +
  scale_fill_manual(values = c('#4E86A7', '#6B9F88', '#F1C04A', '#F07F47', '#D08690', '#9E7DB4')) +
  scale_color_manual(values = c('#4E86A7', '#6B9F88', '#F1C04A', '#F07F47', '#D08690', '#9E7DB4')) +
  scale_y_continuous(limits = c(0, 15)) +
  guides(
    color = guide_legend(nrow = 2),
    fill = guide_legend(nrow = 2)
  ) +
  theme(
    text = element_text(size = 10),
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    legend.position = "bottom",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9))

ggsave("/Users/genesisrodriguez/Downloads/Statistical Analysis/Outputs/Graphs/diagnosisyear_graph.tif", plot = f_diagnosisyear, width = 10, height = 8, dpi = 300,  bg = "white")



```

```{r grouped_diagnosisyear_table}

 pr_smr_g_diagnosisyear <- pr_smr_diagnosisyear %>%
  mutate(`Year of Diagnosis` = case_when(
    `Year of Diagnosis` >= 2004 & `Year of Diagnosis` <= 2009 ~ "2004-2009",
    `Year of Diagnosis` >= 2010 & `Year of Diagnosis` <= 2015 ~ "2010-2015",
    `Year of Diagnosis` >= 2016 & `Year of Diagnosis` <= 2021 ~ "2016-2021"
  )) %>%
  group_by(race_eth, `Year of Diagnosis`) %>%
  summarise(
    Observed = sum(Observed),
    Expected = sum(Expected),
    .groups = "drop"
  ) %>%
  mutate(SMR = Observed / Expected,
         CI.Lower = ifelse(
           Observed == 0, 0, 
           qchisq(0.05 / 2, df = 2 * Observed) / (2 * Expected)
         ),
         CI.Upper =  ifelse(
           Observed == 0, Inf, 
           qchisq(1 - 0.05 / 2, df = 2 * (Observed + 1)) / (2 * Expected))
         )

# US SMRs by Death Year
us_smr_g_diagnosisyear <- us_smr_diagnosisyear %>%
  mutate(`Year of Diagnosis` = case_when(
   `Year of Diagnosis` >= 2004 & `Year of Diagnosis` <= 2009 ~ "2004-2009",
    `Year of Diagnosis` >= 2010 & `Year of Diagnosis` <= 2015 ~ "2010-2015",
    `Year of Diagnosis` >= 2016 & `Year of Diagnosis` <= 2021 ~ "2016-2021"
  ),
  Observed = as.numeric(Observed),
  Expected = as.numeric(Expected)) %>%
  group_by(race_eth, `Year of Diagnosis`) %>%
   summarise(
    Observed = sum(Observed),
    Expected = sum(Expected),
    .groups = "drop") %>%
  mutate(SMR = Observed / Expected,
         CI.Lower = ifelse(
           Observed == 0, 0, 
           SMR * exp(-1.96 * sqrt(1 / Observed))
         ),
         CI.Upper =  ifelse(
           Observed == 0, Inf, 
           SMR * exp(1.96 * sqrt(1 / Observed)))) %>%
  arrange(race_eth, `Year of Diagnosis`)

# Merging PR and US by Diagnosis Year 

merged_g_diagnosisyear <- rbind(pr_smr_g_diagnosisyear, us_smr_g_diagnosisyear) %>%
  mutate(`Year of Diagnosis` = as.factor(`Year of Diagnosis`),
         race_eth = factor(race_eth, levels = c("PR Hispanic", "US Hispanic", "White", "Black", "Asian American, Native Hawaiian and other Pacific Islanders")),
         across(c(Expected:CI.Upper), ~ format(round(as.numeric(.), 1), nsmall = 1))) %>%
  arrange(race_eth)

# Suppressing values under 10

merged_g_diagnosisyear <- suppress_values(merged_g_diagnosisyear)


# Renaming other race/ethnicity

merged_g_diagnosisyear <- recode_race_eth(merged_g_diagnosisyear)

# Widening by Race/Ethnicity

merged_g_diagnosisyear_wider <- merged_g_diagnosisyear %>%
 mutate("95% CI" = ifelse(Observed != "<10",
         paste0("(", CI.Lower, " - ", CI.Upper, ")"),
         "-")) %>%
  select(-c(CI.Lower, CI.Upper)) %>%
  pivot_wider(names_from = c(race_eth), values_from = c(Observed, Expected, SMR, "95% CI")) |> 
  arrange("Year of Diagnosis")

# Naming columns
desired_colnames <- name_columns(merged_g_diagnosisyear_wider)

# Creating gt table

gt_g_diagnosisyear <- create_gt_table(merged_g_diagnosisyear_wider)

# Saving table
gtsave(gt_diagnosisyear, "~/Downloads/Statistical Analysis/Outputs/Tables/Grouped_DeathYear_table.html")

# webshot::webshot("/Users/genesisrodriguez/Downloads/Statistical Analysis/Outputs/Tables/Grouped_DeathYear_table.html",
#                  path.expand("~/Downloads/Statistical Analysis/Outputs/Tables/Grouped_DeathYear_table.png"),
#                  vwidth = 2450,
#                  vheight = 1300)

```

```{r grouped_diagnosisyear_graph}

f_g_diagnosisyear <- ggplot(merged_g_diagnosisyear, aes(y = `Year of Diagnosis`, x = as.numeric(SMR))) +
    geom_point(size = 0.5, position = position_dodge(width = 0.5)) +
    geom_errorbar(
        aes(xmin = as.numeric(CI.Lower), xmax = as.numeric(CI.Upper)),
        width = 0.2, position = position_dodge(width = 0.5)) +
    geom_vline(xintercept = 1, linetype = "solid", color = "red", size = 0.2) +
      geom_text(aes(label = `SMR`), vjust = -0.5, size = 3) +
    facet_wrap(~ race_eth, ncol = 2) +
    theme_bw() +
    labs(x = "", x = "SMR (95% CI)") +
    theme(
        strip.background = element_rect(fill = NA, color = NA),
        legend.position = "none",
        panel.spacing = unit(0.5, "lines"))


ggsave("/Users/genesisrodriguez/Downloads/Statistical Analysis/Outputs/Graphs/g_deathyear.tif", plot = f_g_diagnosisyear, width = 10, height = 8, dpi = 300,  bg = "white")

```

```{r agegroup_diagnosis_table}

# PR SMRs by Age Group
c <- lexpand(cancer_observed, status = cvdStatus != 0, 
             birth = birth_date,
             entry = FollowUpStart, # Without those that did not meet follow-up
             exit = FollowUpEnd, 
             breaks = list(per = 2004:2022, age = c(0, 40, 50, 60, 70, 80, Inf)),
             aggre = list(agegroup = age, deathyear = per, AgeDxCat))

pr_smr_dx_agegroup <- sir(coh.data = c, coh.obs = 'from0to1', 
                  coh.pyrs = 'pyrs', 
                  ref.data = us_reference, 
                  ref.rate = 'gen_cvd_rate', 
                  adjust = c('deathyear', 'agegroup'), 
                  test.type = "homogeneity",
                  conf.type = "univariate",
                  conf.level = 0.95, print = "AgeDxCat") %>%
  as_tibble() %>%
  mutate(race_eth = "PR Hispanic") %>%
  select(-c(pyrs, p_value)) %>%
  rename(`Age at Diagnosis` = AgeDxCat,
         Observed = observed,
         Expected = expected,
         SMR = sir,
        CI.Lower = sir.lo,
        CI.Upper = sir.hi) %>%
     mutate(CI.Lower = qchisq(0.05 / 2, df = 2 * Observed) / (2 * Expected),
    CI.Upper = qchisq(1 - 0.05 / 2, df = 2 * (Observed + 1)) / (2 * Expected))  %>% 
  relocate(race_eth)

# US Age Groups
us_smr_dx_agegroup <- suppressWarnings(tryCatch({
  read.delim("~/Downloads/Statistical Analysis/Inputs/smr_dx_agegroup.txt")
}, error = function(e) {
read.delim("C:/Users/lcastro/OneDrive - Centro Comprensivo de Cancer UPR/projects_shared/BBC/Project_Dra. Carola/Data/Inputs/smr_dx_agegroup.txt")
})) %>%
select(c(Age.recode.Dx,
  Race.Ethnicity.CVD, 
         Observed, Expected, O.E, CI.Lower, CI.Upper)) %>%
  rename(`Age at Diagnosis` = Age.recode.Dx, 
         race_eth = Race.Ethnicity.CVD,
         SMR = O.E) %>% 
  relocate(race_eth) %>%
  mutate(SMR = gsub("#", "", SMR),
         Observed = gsub(",", "", Observed),
         Expected = gsub(",", "", Expected))
  
# Merging PR and US

merged_dx_agegroup <- rbind(pr_smr_dx_agegroup, us_smr_dx_agegroup) %>%
  mutate(across(c(race_eth, "Age at Diagnosis"), as.factor),
         Observed = as.numeric(Observed),
         across(c(Expected:CI.Upper), ~ format(round(as.numeric(.), 1), nsmall = 1)),
         race_eth = factor(race_eth, levels = c("PR Hispanic", "US Hispanic",
                                                "White", "Black", "Asian American, Native Hawaiian and other Pacific Islanders"))) %>%
  arrange(race_eth)

# Supressing values under 10

merged_dx_agegroup <- suppress_values(merged_dx_agegroup)

# Renaming other race/ethnicity

merged_dx_agegroup <- recode_race_eth(merged_dx_agegroup)

# Merging US and PR by Age Group
merged_dx_agegroup_wider <- merged_dx_agegroup %>%
 mutate("95% CI" = ifelse(Observed != "<10",
         paste0("(", CI.Lower, " - ", CI.Upper, ")"),
         "-")) %>%
  select(-c(CI.Lower, CI.Upper)) %>%
  pivot_wider(names_from = c(race_eth), values_from = c(Observed, Expected, SMR, "95% CI")) |> 
  arrange("Age Group")

# Setting column names
desired_colnames <- name_columns(merged_dx_agegroup_wider)

# Creating gt table
gt_dx_agegroup <- create_gt_table(merged_dx_agegroup_wider)

# Saving table

gtsave(gt_dx_agegroup, "~/Downloads/Statistical Analysis/Outputs/Tables/Dx_AgeGroup_table.html")

```

```{r agegroup_diagnosis_graph}

# Create the ggplot
f_dx_agegroup <- ggplot(merged_dx_agegroup[merged_dx_agegroup$`Age at Diagnosis` != "18-39", ], aes(x = `race_eth`, y = as.numeric(`SMR`), fill = `race_eth`)) +
    geom_bar(stat = "identity") + 
    facet_wrap(~ `Age at Diagnosis`, scales = "fixed") + 
    labs(
        x = "Age at Diagnosis",
        y = "Standardized Mortality Ratio (SMR)",
        fill = "Race/Ethnicity"
    ) +
    theme_minimal() +
    geom_text(aes(label = `SMR`), vjust = -0.5, size = 3) +  # Add SMR value above each bar
    theme(plot.title = element_text(face = "bold", hjust = 0.5),
        axis.text.x = element_blank(),
        strip.text = element_text(size = 10, face = "bold")) +
            scale_fill_manual(values = c('#4E86A7', '#6B9F88', '#F1C04A', '#F07F47', '#D08690', '#9E7DB4'),
                               labels = function(x) str_wrap(x, width = 25)) +
  scale_y_continuous(
        breaks = scales::pretty_breaks(n = 6), limits = c(0, 20))

ggsave("/Users/genesisrodriguez/Downloads/Statistical Analysis/Outputs/Graphs/agegroup_diagnosis_graph.tif", plot = f_dx_agegroup, width = 10, height = 8, dpi = 300,  bg = "white")

```

```{r agegroup_death_table}

# PR SMRs by Age Group
c <- lexpand(cancer_observed, status = cvdStatus != 0, 
             birth = birth_date,
             entry = FollowUpStart, # Without those that did not meet follow-up
             exit = FollowUpEnd, 
             breaks = list(per = 2004:2022, age = c(0, 40, 50, 60, 70, 80, Inf)),
             aggre = list(agegroup = age, deathyear = per))

pr_smr_agegroup <- sir(coh.data = c, coh.obs = 'from0to1', 
                  coh.pyrs = 'pyrs', 
                  ref.data = us_reference, 
                  ref.rate = 'gen_cvd_rate', 
                  adjust = c('deathyear', 'agegroup'), 
                  test.type = "homogeneity",
                  conf.type = "univariate",
                  conf.level = 0.95, print = "agegroup") %>%
  as_tibble() %>%
  mutate(race_eth = "PR Hispanic",
         agegroup = case_when(
           agegroup == 0 ~ "18-39",
           agegroup == 80 ~ "80+",
           .default =  paste0(agegroup, "-", agegroup+9))) %>%
  select(-c(pyrs, p_value)) %>%
  rename("Age Group" = agegroup,
         Observed = observed,
         Expected = expected,
         SMR = sir,
        CI.Lower = sir.lo,
        CI.Upper = sir.hi) %>%
     mutate(CI.Lower = qchisq(0.05 / 2, df = 2 * Observed) / (2 * Expected),
    CI.Upper = qchisq(1 - 0.05 / 2, df = 2 * (Observed + 1)) / (2 * Expected))  %>% 
  relocate(race_eth)

# US Age Groups
us_smr_agegroup <- suppressWarnings(tryCatch({
  read.delim("~/Downloads/Statistical Analysis/Inputs/smr_agegroup.txt")
}, error = function(e) {
read.delim("C:/Users/lcastro/OneDrive - Centro Comprensivo de Cancer UPR/projects_shared/BBC/Project_Dra. Carola/Data/Inputs/smr_agegroup.txt")
  } )) %>%
select(c(Attained.Age,
  Race.Ethnicity.CVD, 
         Observed, Expected, O.E, CI.Lower, CI.Upper)) %>%
  filter(Attained.Age != "Total") %>%
  rename("Age Group" = Attained.Age, 
         race_eth = Race.Ethnicity.CVD,
         SMR = O.E) %>% 
  relocate(race_eth) %>%
  mutate(SMR = gsub("#", "", SMR),
         Observed = gsub(",", "", Observed),
         Expected = gsub(",", "", Expected),
        `Age Group` = gsub("0-39", "18-39", `Age Group`))
  
# Merging PR and US

merged_deathyear_agegroup <- rbind(pr_smr_agegroup, us_smr_agegroup) %>%
  mutate(across(c(race_eth, "Age Group"), as.factor),
         Observed = as.numeric(Observed),
         across(c(Expected:CI.Upper), ~ format(round(as.numeric(.), 1), nsmall = 1)),
         race_eth = factor(race_eth, levels = c("PR Hispanic", "US Hispanic",
                                                "White", "Black", "Asian American, Native Hawaiian and other Pacific Islanders"))) %>%
  arrange(race_eth)

# Supressing values under 10

merged_deathyear_agegroup <- suppress_values(merged_deathyear_agegroup)

# Renaming other race/ethnicity

merged_deathyear_agegroup <- recode_race_eth(merged_deathyear_agegroup)

# Merging US and PR by Age Group
merged_deathyear_agegroup_wider <- merged_deathyear_agegroup %>%
 mutate("95% CI" = ifelse(Observed != "<10",
         paste0("(", CI.Lower, " - ", CI.Upper, ")"),
         "-")) %>%
  select(-c(CI.Lower, CI.Upper)) %>%
  pivot_wider(names_from = c(race_eth), values_from = c(Observed, Expected, SMR, "95% CI")) |> 
  arrange("Age Group")

# Setting column names
desired_colnames <- name_columns(merged_deathyear_agegroup_wider)

# Creating gt table
gt_agegroup <- create_gt_table(merged_deathyear_agegroup_wider)

```

```{r agegroup_death_graph}

# Create the ggplot
f_agegroup <- ggplot(merged_deathyear_agegroup[merged_deathyear_agegroup$`Age Group` != "18-39", ], aes(x = `race_eth`, y = as.numeric(`SMR`), fill = `race_eth`)) +
    geom_bar(stat = "identity") + 
    facet_wrap(~ `Age Group`, scales = "fixed") + 
    labs(
        x = "Age at Death",
        y = "Standardized Mortality Ratio (SMR)",
        fill = "Race/Ethnicity"
    ) +
    theme_minimal() +
    geom_text(aes(label = `SMR`), vjust = -0.5, size = 3) +  # Add SMR value above each bar
    theme(axis.text.x = element_blank(),
        strip.text = element_text(size = 10, face = "bold")) +
            scale_fill_manual(values = c('#4E86A7', '#6B9F88', '#F1C04A', '#F07F47', '#D08690', '#9E7DB4'),
                               labels = function(x) str_wrap(x, width = 25)) +
  scale_y_continuous(
        breaks = scales::pretty_breaks(n = 6), limits = c(0, 20))
  

ggsave("/Users/genesisrodriguez/Downloads/Statistical Analysis/Outputs/Graphs/agegroup_deathyear_graph.tif", plot = f_agegroup, width = 10, height = 8, dpi = 300,  bg = "white")



```

```{r stage_table}

# PR SMRs by Death Year
c <- lexpand(cancer_observed, status = cvdStatus != 0, 
             birth = birth_date,
             entry = FollowUpStart, # Without those that did not meet follow-up
             exit = FollowUpEnd, 
             breaks = list(per = 2004:2022, age = c(0, 40, 50, 60, 70, 80, Inf)),
             aggre = list(agegroup = age, deathyear = per, StageAtDx))

pr_smr_stage <- sir(coh.data = c, coh.obs = 'from0to1', 
                  coh.pyrs = 'pyrs', 
                  ref.data = us_reference, 
                  ref.rate = 'gen_cvd_rate', 
                  adjust = c('deathyear', 'agegroup'), 
                  test.type = "homogeneity",
                  conf.type = "univariate",
                  conf.level = 0.95, print = "StageAtDx") %>%
  as_tibble() %>%
  mutate(race_eth = "PR Hispanic") %>%
  select(-c(pyrs, p_value)) %>%
  rename(Stage = StageAtDx,
         Observed = observed,
         Expected = expected,
         SMR = sir,
        CI.Lower = sir.lo,
        CI.Upper = sir.hi) %>%
   mutate(Stage = ifelse(Stage == "Unstaged", "Unknown/unstaged", Stage),
          CI.Lower = qchisq(0.05 / 2, df = 2 * Observed) / (2 * Expected),
    CI.Upper = qchisq(1 - 0.05 / 2, df = 2 * (Observed + 1)) / (2 * Expected))  %>% 
  relocate(race_eth)

# US SMRs by Stage
us_smr_stage <- suppressWarnings(tryCatch({
  read.delim("~/Downloads/Statistical Analysis/Inputs/smr_stage.txt")
}, error = function(e) {
read.delim("C:/Users/lcastro/OneDrive - Centro Comprensivo de Cancer UPR/projects_shared/BBC/Project_Dra. Carola/Data/Inputs/smr_stage.txt")
  } )) %>%
select(c(Combined.Summary.Stage..2004..,
  Race.Ethnicity.CVD, 
         Observed, Expected, O.E, CI.Lower, CI.Upper)) %>%
  filter(Race.Ethnicity.CVD != "All races/ethnicities",
         !(Combined.Summary.Stage..2004.. %in% c("In situ", "N/A", "Blank(s)"))) %>%
  rename(Stage = Combined.Summary.Stage..2004.., 
         race_eth = Race.Ethnicity.CVD,
         SMR = O.E) %>% 
  relocate(race_eth) %>%
  mutate(SMR = gsub("#", "", SMR),
         Observed = gsub(",", "", Observed),
         Expected = gsub(",", "", Expected))

# Merging PR and US by Death Year 

merged_stage <- rbind(pr_smr_stage, us_smr_stage) %>%
  mutate(across(c(race_eth, Stage), as.factor),
         across(c(Expected:CI.Upper), ~ format(round(as.numeric(.), 1), nsmall = 1)),
         race_eth = factor(race_eth, levels = c("PR Hispanic", "US Hispanic",
                                                "White", "Black", "Asian American, Native Hawaiian and other Pacific Islanders")),
         Stage = factor(Stage,
         levels = c("Localized", "Regional", "Distant", "Unknown/unstaged"))) %>%
  arrange(race_eth, Stage)

# Suppressing values under 10

merged_stage <- suppress_values(merged_stage)

# Renaming other race/ethnicity

merged_stage <- recode_race_eth(merged_stage)

# Suppressing values for Stage = Unknown/unstaged

merged_stage <- merged_stage %>%
    mutate(
      Expected = ifelse(Stage == "Unknown/unstaged", "-", Expected),
      SMR = ifelse(Stage == "Unknown/unstaged", "-", SMR),
      CI.Lower = ifelse(Stage == "Unknown/unstaged", "-", CI.Lower),
      CI.Upper =  ifelse(Stage == "Unknown/unstaged", "-", CI.Upper))

# Widening by Race/Ethnicity

merged_stage_wider <- merged_stage %>%
 mutate("95% CI" = ifelse(Expected != "-",
         paste0("(", CI.Lower, " - ", CI.Upper, ")"),
         "-")) %>%
  select(-c(CI.Lower, CI.Upper)) %>%
  pivot_wider(names_from = c(race_eth), values_from = c(Observed, Expected, SMR, "95% CI"))




# Naming columns
desired_colnames <- name_columns(merged_stage_wider)

# Creating gt table

gt_stage <- create_gt_table(merged_stage_wider)

# Saving table
gtsave(gt_stage, "~/Downloads/Statistical Analysis/Outputs/Tables/Stage_table.html")


```

```{r procedure_table}

# PR SMRs by Death Year
c <- lexpand(cancer_observed, status = cvdStatus != 0, 
             birth = birth_date,
             entry = FollowUpStart, # Without those that did not meet follow-up
             exit = FollowUpEnd, 
             breaks = list(per = 2004:2022, age = c(0, 40, 50, 60, 70, 80, Inf)),
             aggre = list(agegroup = age, deathyear = per, Procedure))

pr_smr_procedure <- sir(coh.data = c, coh.obs = 'from0to1', 
                  coh.pyrs = 'pyrs', 
                  ref.data = us_reference, 
                  ref.rate = 'gen_cvd_rate', 
                  adjust = c('deathyear', 'agegroup'), 
                  test.type = "homogeneity",
                  conf.type = "univariate",
                  conf.level = 0.95, print = "Procedure") %>%
  as_tibble() %>%
  mutate(race_eth = "PR Hispanic") %>%
  select(-c(pyrs, p_value)) %>%
  rename(Treatment = Procedure,
         Observed = observed,
         Expected = expected,
         SMR = sir,
        CI.Lower = sir.lo,
        CI.Upper = sir.hi) %>%
   mutate(CI.Lower = qchisq(0.05 / 2, df = 2 * Observed) / (2 * Expected),
    CI.Upper = qchisq(1 - 0.05 / 2, df = 2 * (Observed + 1)) / (2 * Expected))  %>% 
  relocate(race_eth)

# US SMRs by Procedure
us_smr_procedure <- suppressWarnings(tryCatch({
  read.delim("~/Downloads/Statistical Analysis/Inputs/smr_procedure.txt")
}, error = function(e) {
read.delim("C:/Users/lcastro/OneDrive - Centro Comprensivo de Cancer UPR/projects_shared/BBC/Project_Dra. Carola/Data/Inputs/smr_procedure.txt")
  } ))  %>%
select(c(Merged.Procedures,
  Race.Ethnicity.CVD, 
         Observed, Expected, O.E, CI.Lower, CI.Upper)) %>%
  filter(Race.Ethnicity.CVD != "All races/ethnicities") %>%
  rename(Treatment = Merged.Procedures,
         race_eth = Race.Ethnicity.CVD,
         SMR = O.E) %>% 
  relocate(race_eth) %>%
  mutate(SMR = gsub("#", "", SMR),
         Observed = gsub(",", "", Observed),
         Expected = gsub(",", "", Expected),
         Treatment = gsub("\\[OTHER\\]", "Other", Treatment, ignore.case = FALSE))

# Merging PR and US by Death Year 

merged_procedure <- rbind(pr_smr_procedure, us_smr_procedure) %>%
  mutate(across(c(race_eth, Treatment), as.factor),
         across(c(Expected:CI.Upper), ~ format(round(as.numeric(.), 1), nsmall = 1)),
         race_eth = factor(race_eth, levels = c("PR Hispanic", "US Hispanic",
                                                "White", "Black", "Asian American, Native Hawaiian and other Pacific Islanders")),
         Treatment = factor(Treatment,
                               levels = c("Surgery Only", "Chemotherapy Only",
                                          "Radiotherapy Only", "Chemotherapy and Radiotherapy",
                                          "Other"))) %>%
  arrange(race_eth, Treatment)

# Suppressing values under 10

merged_procedure <- suppress_values(merged_procedure)

# Renaming other race/ethnicity

merged_procedure <- recode_race_eth(merged_procedure)

# Widening by Race/Ethnicity

merged_procedure_wider <- merged_procedure %>%
 mutate("95% CI" = ifelse(Observed != "<10",
         paste0("(", CI.Lower, " - ", CI.Upper, ")"),
         "-")) %>%
  select(-c(CI.Lower, CI.Upper)) %>%
  pivot_wider(names_from = c(race_eth), values_from = c(Observed, Expected, SMR, "95% CI"))

# Naming columns
desired_colnames <- name_columns(merged_procedure_wider)

# Creating gt table

gt_procedure <- create_gt_table(merged_procedure_wider)

# Saving table
gtsave(gt_procedure, "~/Downloads/Statistical Analysis/Outputs/Tables/Procedure_table.html")

# webshot::webshot("/Users/genesisrodriguez/Downloads/Statistical Analysis/Outputs/Tables/Procedure_table.html",
#                  path.expand("~/Downloads/Statistical Analysis/Outputs/Tables/Procedure_table.png"),
#                  vwidth = 2450,
#                  vheight = 1300)

```

```{r proc_stage_table, results='asis'}

# PR SMRs - Numerator
c <- lexpand(cancer_observed, status = cvdStatus != 0, 
             birth = birth_date,
             entry = FollowUpStart, # Without those that did not meet follow-up
             exit = FollowUpEnd, 
             breaks = list(per = 2004:2022, age = c(0, 40, 50, 60, 70, 80, Inf)),
             aggre = list(agegroup = age, deathyear = per, Procedure = Procedure, StageAtDx))

# PR SMRs by Procedure and Stage

pr_smr_proc_stage <- sir(coh.data = c, coh.obs = 'from0to1', 
                  coh.pyrs = 'pyrs', 
                  ref.data = us_reference, 
                  ref.rate = 'gen_cvd_rate', 
                  adjust = c('deathyear', 'agegroup'), 
                  test.type = "homogeneity",
                  conf.type = "univariate",
                  conf.level = 0.95, print = list(Procedure, StageAtDx)) %>%
  as_tibble() %>%
  mutate(race_eth = "PR Hispanic",
         StageAtDx = ifelse(StageAtDx == "Unstaged", "Unknown/unstaged", StageAtDx)) %>%
  select(-c(pyrs, p_value)) %>%
  rename(Stage = StageAtDx,
         Observed = observed,
         Expected = expected,
         SMR = sir,
        CI.Lower = sir.lo,
        CI.Upper = sir.hi) %>%
   mutate(CI.Lower = qchisq(0.05 / 2, df = 2 * Observed) / (2 * Expected),
    CI.Upper = qchisq(1 - 0.05 / 2, df = 2 * (Observed + 1)) / (2 * Expected))  %>% 
  relocate(race_eth)

# US SMRs by Procedure and Stage

us_smr_proc_stage <- suppressWarnings(tryCatch({
  read.delim("~/Downloads/Statistical Analysis/Inputs/smr_procedure_stage.txt")
}, error = function(e) {
read.delim("C:/Users/lcastro/OneDrive - Centro Comprensivo de Cancer UPR/projects_shared/BBC/Project_Dra. Carola/Data/Inputs/smr_procedure_stage.txt")
  } )) %>%
  filter(!(Combined.Summary.Stage..2004.. %in% c("In situ", "Blank(s)", "N/A")),
         Race.Ethnicity.CVD != "All races/ethnicities") %>%
select(c(Merged.Procedures, Combined.Summary.Stage..2004.., Race.Ethnicity.CVD, 
         Observed, Expected, O.E, CI.Lower, CI.Upper)) %>%
  rename(Procedure = Merged.Procedures, 
         Stage = Combined.Summary.Stage..2004.., 
         race_eth = Race.Ethnicity.CVD,
         SMR = O.E) %>% 
  relocate(race_eth, Procedure) %>%
  mutate(Procedure = ifelse(Procedure == "[OTHER]", "Other", Procedure),
         SMR = gsub("#", "", SMR),
         Observed = gsub(",", "", Observed),
         Expected = gsub(",", "", Expected),
         race_eth = gsub("Asian American, Native Hawaiian and other Pacific Islanders", "AANHPI & AIAN", race_eth))
  
# US and PR Procedure and Stage

merged_proc_stage <- rbind(pr_smr_proc_stage, us_smr_proc_stage) %>%
    mutate(Procedure = factor(Procedure,
                               levels = c("Surgery Only", "Chemotherapy Only",
                                          "Radiotherapy Only", "Chemotherapy and Radiotherapy",
                                          "Other")),
         Stage = factor(Stage, levels = c("Localized", "Regional", "Distant", "Unknown/unstaged")),
         race_eth = factor(race_eth, levels = c("PR Hispanic", "US Hispanic",
                                                "White", "Black", "AANHPI & AIAN"))) %>%
  arrange(race_eth)

# Supressing values under 10

merged_proc_stage <- suppress_values(merged_proc_stage)

# Suppressing values for Stage = Unknown/unstaged

merged_proc_stage <- merged_proc_stage %>%
    mutate(
      Expected = ifelse(Procedure == "Other", "-", Expected),
      SMR = ifelse(Procedure == "Other", "-", SMR),
      CI.Lower = ifelse(Procedure == "Other", "-", CI.Lower),
      CI.Upper =  ifelse(Procedure == "Other", "-", CI.Upper))

# Widen table based on Race/Ethnicity

merged_proc_stage_wider <- merged_proc_stage %>%
 mutate("95% CI" = ifelse(Expected != "-",
         paste0("(", CI.Lower, " - ", CI.Upper, ")"),
         "-")) %>%
  select(-c(CI.Lower, CI.Upper)) %>%
  filter(Stage != "Unknown/unstaged") %>%
  pivot_wider(names_from = c(race_eth), values_from = c(Observed, Expected, SMR, "95% CI")) |> 
  mutate(Procedure = factor(Procedure,
                               levels = c("Surgery Only", "Chemotherapy Only",
                                          "Radiotherapy Only", "Chemotherapy and Radiotherapy",
                                          "Other")),
         Stage = factor(Stage, levels = c("Localized", "Regional", "Distant"))) |>
  arrange(Procedure, Stage)

# Column titles
desired_colnames <- name_columns(merged_proc_stage_wider)

# Creating gt table
gt_proc_stage <- create_gt_table(merged_proc_stage_wider, groupname_col = "Procedure")

# Printing
gtsave(gt_proc_stage, "~/Downloads/Statistical Analysis/Outputs/Tables/Supplementary_Table_4.html")

```

```{r proc_stage_figure}

merged_proc_stage2 <- merged_proc_stage %>%
  filter(Stage != "Unknown/unstaged") %>%
  mutate(race_eth = case_when(
    race_eth == "Asian American, Native Hawaiian and other Pacific Islanders" ~ "Asian American,\nNative Hawaiian,\nand other Pacific Islanders",
    TRUE ~ race_eth),
    race_eth = factor(race_eth, levels = c("PR Hispanic", "US Hispanic",
               "White", "Black", "AANHPI & AIAN")),
     Procedure = str_wrap(Procedure, width = 15))

f_proc_stage <- ggplot(merged_proc_stage2[merged_proc_stage2$Procedure != "Other", ], aes(y = Stage, x = as.numeric(SMR))) +
    geom_point(size = 0.5, position = position_dodge(width = 0.5)) +
    geom_errorbar(
        aes(xmin = as.numeric(CI.Lower), xmax = as.numeric(CI.Upper)),
        width = 0.2, position = position_dodge(width = 0.5)
    ) +
    geom_vline(xintercept = 1, linetype = "solid", color = "red", size = 0.2) +
    facet_grid(Procedure ~ race_eth, scales = "fixed", space = "fixed", switch = "y") +
   geom_text(aes(label = round(as.numeric(SMR), 2)), 
            vjust = -1.7, size = 3, fontface = "bold", color = "grey30") +
  theme_minimal() +
    labs(y = "", x = "SMR (95% CI)") +
    theme(
      #  text = element_text(size = 7), 
        strip.text.y = element_text(angle = 0, face = "bold"), #size = 6), 
       # strip.text.x = element_text(size = 6), 
        strip.background = element_rect(fill = NA, color = NA),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(hjust = 1),  
       # axis.text.y = element_text(size = 6), 
        legend.position = "none",
        panel.spacing.x = unit(1, "lines"),
        panel.spacing.y = unit(1, "lines"),
        panel.grid = element_blank()) +
  scale_x_continuous(limits = c(-5,10), breaks = c(-5, 1, 5.0, 10))

ggsave("/Users/genesisrodriguez/Downloads/Statistical Analysis/Outputs/Graphs/proc_stage_graph.tif", plot = f_proc_stage, width = 8, height = 6.5, dpi = 1200,  bg = "white")

```

```{r smr_table1}

table1_general <- merged_general_wider %>%
  mutate(Variable = "Overall") %>%
         #across(starts_with("Observed_"), ~ round(as.numeric(.x), 0))) %>%
  relocate(Variable)

table1_agegroup <- merged_deathyear_agegroup_wider %>%
  rename(Variable = `Age Group`) %>%
  bind_rows(tibble(Variable = "Age at Death"),.) %>%
  relocate(Variable)

 table1_dx_agegroup <- merged_dx_agegroup_wider %>%
  rename(Variable = `Age at Diagnosis`) %>%
  bind_rows(tibble(Variable = "Age at Diagnosis"),.) %>%
  relocate(Variable)
 
 table1_stage <- merged_stage_wider %>%
  rename(Variable = Stage) %>%
  bind_rows(tibble(Variable = "Stage"),.) %>%
  relocate(Variable)
 
 table1_procedure <- merged_procedure_wider %>%
  rename(Variable = Treatment) %>%
  bind_rows(tibble(Variable = "Treatment"),.) %>%
  relocate(Variable)
 
 table1_g_deathyear <- merged_g_deathyear_wider %>%
  rename(Variable = `Year of Death`) %>%
  bind_rows(tibble(Variable = "Year of Death"),.) %>%
  relocate(Variable)

 table1_g_diagnosisyear <- merged_g_diagnosisyear_wider %>%
  rename(Variable = `Year of Diagnosis`) %>%
  bind_rows(tibble(Variable = "Year of Diagnosis"),.) %>%
  relocate(Variable)
 
combined_table <- rbind(table1_general, table1_dx_agegroup, table1_stage, table1_procedure, table1_g_diagnosisyear, table1_g_deathyear) %>%
    mutate(has_na = apply(., 1, function(row) any(is.na(row))), # Rows to make bold
           across(everything(), ~ ifelse(is.na(.), "", .))) # Make NAs Blank
 
desired_colnames <- name_columns(combined_table)

gt_smr_table1 <- combined_table %>%
   create_gt_table() %>%
    tab_style(
        style = cell_text(weight = "bold"),
        locations = cells_body(rows = has_na)) %>%
  cols_hide(columns = has_na)

```

```{r sir}

# SIR session

pr_sir_general <- cancer_observed %>%
  mutate(Vital = as.character(Vital),
    `Cause of Death` = case_when(
      Vital == "CVD death" ~ "CVD Death",
      Vital == "BC death" ~ "Other Causes of Death",
      Vital == "Other Death" ~ "Other Causes of Death",
      TRUE ~ Vital)) %>%
  group_by(`Cause of Death`) %>%
  count(name = "Count") %>%
  mutate(race_eth = "PR Hispanic")

us_sir_general <- suppressWarnings(tryCatch({
  read.delim("~/Downloads/Statistical Analysis/Inputs/")
}, error = function(e) {
read.delim("C:/Users/lcastro/OneDrive - Centro Comprensivo de Cancer UPR/projects_shared/BBC/Project_Dra. Carola/Data/Inputs/sir_general.txt")
  } )) %>%
select(Index.Record.Cause.of.Death.CVD.recode,
       Index.Record.Race.Ethnicity.CVD, 
       Count) %>%
  rename(`Cause of Death` = Index.Record.Cause.of.Death.CVD.recode,
         race_eth = Index.Record.Race.Ethnicity.CVD) %>% 
  relocate(race_eth) %>%
  mutate(Count = as.numeric(gsub(",", "", Count)),
        `Cause of Death` = gsub("\\[OTHER\\]", "Other Causes of Death",
    `Cause of Death`),
    `Cause of Death` = gsub("CVD", "CVD Death",
    `Cause of Death`))

us_sir_general <- recode_race_eth(us_sir_general)

us_sir_general_wider <- us_sir_general %>%
 pivot_wider(names_from = c(race_eth), values_from = Count)


merged_sir_general <- rbind(pr_sir_general, us_sir_general) %>%
  mutate(race_eth = factor(race_eth, levels = c("PR Hispanic", "US Hispanic",
                                                "White", "Black", "AANHPI & AIAN"))) %>%
  group_by(race_eth) %>%
  mutate(Count = paste0(comma(Count), " (", sprintf("%.2f", Count/ sum(Count) * 100), "%)")) %>%
  arrange(race_eth) 

merged_sir_general_wider <- merged_sir_general %>%
  pivot_wider(names_from = c(race_eth), values_from = Count)

# Creating gt table
  
gt_sir_general <- merged_sir_general_wider |> 
  rename(`Vital status` = `Cause of Death`) |>
  gt() |> 
    tab_style(
      style = cell_text(weight = "bold"), 
      locations = cells_column_labels(everything())) |> 
  cols_align(align = "center", columns = everything()); gt_sir_general

>>>>>>> 1fd5aa90eaed4b1c061ae7e268eebeccf08ea326

```

```{r cumulative_mortality}
# cuminc_fit <- cuminc(FollowUpYears, fstatus = cvdStatus, group = cancer_observed)

# km_fit <- survfit(Surv(FollowUpYears, cvdStatus) ~ 1, data = cancer_observed)
# km_fit_2 <- survfit(Surv(FollowUpYears, cvdStatus) ~ Race.Ethnicity.CVD, data = us_sir_general_cl_df)
# 
# # Define time points
# time_points <- c(1, 5, 10)
# 
# # Function to extract Kaplan-Meier estimates
# extract_km <- function(km_fit, race_ethnicity) {
#   summary_km <- summary(km_fit, times = time_points)
#   data.frame(
#     Time = summary_km$time,
#     Cumulative_Mortality = round((1 - summary_km$surv) * 100,2),  # Convert to %
#     Lower_CI = (1 - summary_km$upper) * 100,  # Invert CI
#     Upper_CI = (1 - summary_km$lower) * 100,
#     Race_Ethnicity = race_ethnicity
#   ) %>%
#     mutate(Cumulative_Mortality = sprintf("%.2f%% (%.2f%% to %.2f%%)", Cumulative_Mortality, Lower_CI, Upper_CI))
# }
# 
# # Extract estimates for PR and US racial/ethnic groups
# pr_df <- extract_km(km_fit, "PR Hispanic")
# us_df <- extract_km(km_fit_2, rep(c("AANHPI & AIAN", "Black", "US Hispanic", "White"), each = length(time_points)))
# 
# # Combine and reshape data
# final_table <- bind_rows(pr_df, us_df) %>%
#   select(Race_Ethnicity, Time, Cumulative_Mortality) %>%
#   pivot_wider(names_from = Time, values_from = Cumulative_Mortality, names_prefix = "Year_") %>%
#   rename(
#     `1-year CM (95% CI)` = Year_1,
#     `5-year CM (95% CI)` = Year_5,
#     `10-year CM (95% CI)` = Year_10) %>%
#   mutate(Race_Ethnicity = factor(Race_Ethnicity, levels = c("PR Hispanic", "US Hispanic", "White", "Black", "AANHPI & AIAN"))) %>%
#   arrange(Race_Ethnicity) 
# 
# # Generate gt table
# km_table <- final_table %>%
#   gt(rowname_col = "Race_Ethnicity") %>%
#   tab_header(
#     title = "CVD Cumulative Mortality (CM) and 95% Confidence Intervals",
#     subtitle = "By Race/Ethnicity"
#   ) %>%
#   cols_label(
#     `1-year CM (95% CI)` = "1-year CM (95% CI)",
#     `5-year CM (95% CI)` = "5-year CM (95% CI)",
#     `10-year CM (95% CI)` = "10-year CM (95% CI)"
#   ) %>%
#   tab_options(
#     table.font.size = px(14),
#     column_labels.font.weight = "bold",
#     heading.title.font.size = px(18),
#     heading.subtitle.font.size = px(14),
#     table.border.top.color = "black",
#     table.border.bottom.color = "black"
#   ) %>%
#   tab_style(
#     style = cell_text(weight = "bold"),
#     locations = cells_row_groups()
#   ); km_table

# library(cmprsk)
# cancer_observed$status2 <- case_when(cancer_observed$Vital == "Alive" ~ 0,
#                                 cancer_observed$Vital == "CVD death" ~ 1,
#                                 TRUE ~ 2)
# cm_fit <- crr(cancer_observed$FollowUpYears, cancer_observed$status2, cov1 = cancer_observed[, c("AgeDxCat2", "StageAtDx2", "yearDiagCat2")])
# 
# cuminc(ftime = cancer_observed$FollowUpYears, 
#        fstatus = cancer_observed$status2, 
#        group = cancer_observed$AgeDxCat2)

```

```{r quality_control, include = FALSE}

# Verifying categorization is donde correctly in grid for rates

# SIR Function's
  
c <- lexpand(cancer_observed, status = cvdStatus != 0, 
             birth = birth_date,
             entry = FollowUpStart, # Without those that did not meet follow-up
             exit = FollowUpEnd, 
             breaks = list(per = 2004:2022, age = c(0, 40, 50, 60, 70, 80, Inf)),
             aggre = list(agegroup = age, deathyear = per))

t1 <- c %>% arrange(deathyear, agegroup)

# Manually calculated with wrangling_cancer_registry_patients.Rmd

t2 <- read.csv("~/Downloads/Registro Central de Cancer/cvm_by_year_agegroup2.csv") %>%
  filter(AgeGroup2 != "All ages") %>%
  mutate(AgeGroup2 = as.numeric(substr(AgeGroup2, 1, 2)),
         Year = as.numeric(Year),
         cvd_deaths_BC = ifelse(is.na(cvd_deaths_BC), 0, cvd_deaths_BC)) %>%
  rename(agegroup = AgeGroup2, 
         deathyear = Year)

# Merging

merged_data <- t1 %>%
  left_join(t2, by = c("agegroup", "deathyear")) %>%
  select(-c(from0to0, at.risk, cvd_rate)) %>%
  relocate(PersonYears, .before = pyrs) %>%
  mutate(across(cvd_deaths_BC, ~ replace_na(., 0))) %>%
  bind_rows(summarise(., across(where(is.numeric), sum, na.rm = TRUE))) 

# 2. Checking n for each race/ethnicity is consistent across variables

sum_observed_columns(merged_general_wider)
sum_observed_columns(merged_deathyear_wider)
sum_observed_columns(merged_g_deathyear_wider)
sum_observed_columns(merged_diagnosisyear_wider)
sum_observed_columns(merged_dx_agegroup_wider)
sum_observed_columns(merged_stage_wider)
sum_observed_columns(merged_procedure_wider)

# Race/Ethnicity percentages
rbind(pr_sir_general, us_sir_general) %>% group_by(race_eth) %>% summarise(n = sum(Count)) %>% mutate(perc = n/sum(n)*100)

# Follow-up time
bind_rows(
  us_sir_general_cl_df %>%
    mutate(FollowUpYears = ifelse(FollowUpYears < 0, 0, FollowUpYears)) %>%
    group_by(Race.Ethnicity.CVD) %>%
    summarize(
      Min. = min(FollowUpYears, na.rm = TRUE),
      `1st Qu.` = quantile(FollowUpYears, 0.25, na.rm = TRUE),
      Median = median(FollowUpYears, na.rm = TRUE),
      Mean = mean(FollowUpYears, na.rm = TRUE),
      `3rd Qu.` = quantile(FollowUpYears, 0.75, na.rm = TRUE),
      Max. = max(FollowUpYears, na.rm = TRUE)
    ),
  cancer_observed %>%
    summarize(
      Race.Ethnicity.CVD = "PR Hispanic",
      Min. = min(FollowUpYears, na.rm = TRUE),
      `1st Qu.` = quantile(FollowUpYears, 0.25, na.rm = TRUE),
      Median = median(FollowUpYears, na.rm = TRUE),
      Mean = mean(FollowUpYears, na.rm = TRUE),
      `3rd Qu.` = quantile(FollowUpYears, 0.75, na.rm = TRUE),
      Max. = max(FollowUpYears, na.rm = TRUE)
    )) %>%
  mutate(Race.Ethnicity.CVD = factor(Race.Ethnicity.CVD, levels = c("PR Hispanic", "US Hispanic",
                                                "White", "Black", "Asian American, Native Hawaiian and other Pacific Islanders"))) %>%
  arrange(Race.Ethnicity.CVD)


# Year of diagnosis
us_sir_general_cl_df %>% count(Year.of.diagnosis) %>% 
  mutate(perc = n/sum(n)*100)
cancer_observed %>% count(ydiag) %>% 
  mutate(perc = n/sum(n)*100)

# Year of death
us_sir_general_cl_df %>% 
  count(Year.of.death.recode) %>% 
  mutate(perc = n/sum(n)*100)
cancer_observed %>%
  mutate(Year.of.death.recode = ifelse(FollowUpEnd == as.Date("2021-12-31") | Vital == "Alive", 
                                       "Alive at last contact", 
                                       as.character(year(FollowUpEnd)))) %>%
  count(Year.of.death.recode) %>% 
  mutate(perc = n/sum(n)*100)


# Vital status
us_sir_general_cl_df %>% count(Cause.of.Death.CVD.recode) %>% mutate(perc = n/sum(n))
cancer_observed %>% mutate(Vital = as.character(Vital),
    Vital = case_when(
      Vital == "CVD death" ~ "CVD Death",
      Vital == "BC death" ~ "Other Causes of Death",
      Vital == "Other Death" ~ "Other Causes of Death",
      TRUE ~ Vital)) %>%
  count(Vital) %>% 
  mutate(perc = n/sum(n))

# Percentage of deaths by race/ethnicity

merged_general %>% mutate(Observed = as.numeric(gsub(",", "", Observed)), Percentage_Deaths = Observed/sum(Observed)* 100)

# Mean age at diagnosis

summary(cancer_observed$AgeDx)
summary(cancer_observed$FollowUpYears)

tblFun <- function(x){
    tbl <- table(x)
    res <- cbind(tbl,round(prop.table(tbl)*100,1))
    colnames(res) <- c('Count','Percentage')
    res
}
tblFun(cancer_observed$Procedure)
tblFun(cancer_observed$StageAtDx)
tblFun(cancer_observed$yearDiagCat)
tblFun(cancer_observed$deathYearCat)
tblFun(cancer_observed$Vital)

# Number of deaths
sum(cancer_observed$Vital != "Alive")

# Proportion of CVD deaths
round(sum(cancer_observed$Vital == "CVD death") / sum(cancer_observed$Vital != "Alive") * 100, 2)

```

```{r outline, cache = FALSE, results = 'asis'}

# Flow diagram
f <- 0
cat(paste("\n\n Figure", f + 1, "  Flow diagram defining selection of women diagnosed with first breast cancer between 2004-2019 and followed through 2021 in PR and 17 SEER registries. \n"))
f <- f + 1

image_read("~/Downloads/Statistical Analysis/Outputs/Graphs/Flowchart.tif")

cat("\n\n\\pagebreak\n")

# PR Table 1
t <- 0
cat(paste("\n Table", t + 1, " Baseline Characteristics of women who died in Puerto Rico diagnosed with Breast Cancer from 2004-2019 and followed through 2021 \n"))
t <- t + 1

gt_pr_table1

cat("\n\n\\pagebreak\n")

cat("\n## CVD SMR for PR BC Survivors with PR Population as Referent\n")

gt_pr_general

# PR SMR by deathyear
cat(paste("\n Table", t + 1, " CVD SMRs by Year of Death for BC Survivors in Puerto Rico Using PR Population as Referent\n"))
t <- t + 1

smr_deathyear %>% gt()

# PR SMRs by year of death figure
cat(paste("\n Figure", f + 1, " CVD SMRs by Year of Death for BC survivors in Puerto Rico Using PR population as Referent\n"))
f <- f + 1

f_prdeathyear

cat("\n\n\\pagebreak\n")

# PR SMR by deathyear
cat(paste("\n Table", t + 1, " CVD SMRs by Year of Death for BC Survivors in Puerto Rico Using PR Population as Referent\n"))
t <- t + 1

smr_deathYearCat %>% gt()


# PR SMRs by year of death figure
cat(paste("\n Figure", f + 1, " CVD SMRs by Year of Death for BC survivors in Puerto Rico Using PR population as Referent\n"))
f <- f + 1

f_deathYearCat

cat("\n\n\\pagebreak\n")

# PR SMR by diagnosis year
cat(paste("\n Table", t + 1, " CVD SMRs by Year of Diagnosis for BC Survivors in Puerto Rico Using PR Population as Referent\n"))
t <- t + 1

smr_ydiag %>% gt()


# PR SMRs by year of diagnosis figure
cat(paste("\n Figure", f + 1, " CVD SMRs by Year of Diagnosis for BC survivors in Puerto Rico Using PR population as Referent\n"))
f <- f + 1

f_ydiag

# PR SMRs by year of diagnosis figure
cat(paste("\n V2: Figure", f + 1, " CVD SMRs by Year of Diagnosis for BC survivors in Puerto Rico Using PR population as Referent\n"))
f <- f + 1

f_ydiag2

cat("\n\n\\pagebreak\n")

# PR SMR by diagnosis year
cat(paste("\n Table", t + 1, " CVD SMRs by Year of Diagnosis for BC Survivors in Puerto Rico Using PR Population as Referent\n"))
t <- t + 1

smr_diagCat %>% gt()

cat("\n\n\\pagebreak\n")

# PR SMRs by year of diagnosis figure
cat(paste("\n Figure", f + 1, " CVD SMRs by Year of Diagnosis for BC survivors in Puerto Rico Using PR population as Referent\n"))
f <- f + 1

f_ydiagcat

cat("\n\n\\pagebreak\n")

# PR SMR by agegroup
cat(paste("\n Table", t + 1, " CVD SMRs by Age for BC Survivors in Puerto Rico Using PR Population as Referent\n"))
t <- t + 1

smr_agegroup %>% gt()


# PR SMRs by agegroup figure
cat(paste("\n Figure", f + 1, " CVD SMRs by Age for BC survivors in Puerto Rico Using PR population as Referent\n"))
f <- f + 1

f_agegroup

cat("\n\n\\pagebreak\n")

# PR SMR by treatment
cat(paste("\n Table", t + 1, " CVD SMRs by Initial Treatment Type for BC Survivors in Puerto Rico Using PR Population as Referent\n"))
t <- t + 1

smr_treatment %>% gt()

# PR SMRs by treatment figure
cat(paste("\n Figure", f + 1, " CVD SMRs by Initial Treatment Type for BC survivors in Puerto Rico Using PR population as Referent\n"))
f <- f + 1

f_treatment

cat("\n\n\\pagebreak\n")

# Overall SMR
cat(paste("\n Table", t + 1, " CVD SMRs for BC Survivors in Puerto Rico and US Race/Ethnicity Groups Using US Population as Referent\n"))
t <- t + 1

gt_general


# Overall SMR figure
f <- 0
cat(paste("\n Figure", f + 1, " CVD SMRs for BC survivors in Puerto Rico and US Race/Ethnicity groups using US population as referent\n"))
f <- f + 1
f_general

cat("\n\n\\pagebreak\n")

# SMRs by year of death
cat(paste("\n Table", t + 1, " CVD SMRs by Year of Death for BC Survivors in Puerto Rico and US Race/Ethnicity Groups Using US population as Referent\n"))
t <- t + 1

gt_deathyear

cat("\n\n\\pagebreak\n")

# SMRs by year of death figure
cat(paste("\n Figure", f + 1, " CVD SMRs by Year of Death for BC survivors in Puerto Rico and US Race/Ethnicity Groups Using US population as Referent\n"))
f <- f + 1

f_deathyear

cat("\n\n\\pagebreak\n")

# SMRs by grouped year of death
cat(paste("\n Table", t + 1, " CVD SMRs by Year of Death for BC Survivors in Puerto Rico and US Race/Ethnicity Groups Using US population as Referent\n"))
t <- t + 1

gt_g_deathyear

cat("\n\n\\pagebreak\n")

# SMRs by grouped year of death figure

cat(paste("\n Figure", f + 1, " CVD SMRs by Year of Death for BC survivors in Puerto Rico and US Race/Ethnicity Groups Using US population as Referent\n"))
f <- f + 1
f_g_deathyear

cat("\n\n\\pagebreak\n")

# SMRs by year of diagnosis
cat(paste("\n Table", t + 1, " CVD SMRs by Year of Diagnosis for BC Survivors in Puerto Rico and US Race/Ethnicity Groups Using US population as Referent\n"))
t <- t + 1

gt_diagnosisyear

cat("\n\n\\pagebreak\n")

# SMRs by year of diagnosis figure
cat(paste("\n Figure", f + 1, " CVD SMRs by Year of Diagnosis for BC survivors in Puerto Rico and US Race/Ethnicity Groups Using US population as Referent\n"))
f <- f + 1

f_diagnosisyear

cat("\n\n\\pagebreak\n")

# SMRs by grouped year of diagnosis

cat(paste("\n Table", t + 1, " CVD SMRs by Year of Diagnosis for BC Survivors in Puerto Rico and US Race/Ethnicity Groups Using US population as Referent\n"))
t <- t + 1

gt_g_diagnosisyear

cat("\n\n\\pagebreak\n")

# SMRs by grouped year of diagnosis figure
cat(paste("\n Figure", f + 1, " CVD SMRs by Year of Diagnosis for BC survivors in Puerto Rico and US Race/Ethnicity Groups Using US population as Referent\n"))
f <- f + 1

f_g_diagnosisyear

cat("\n\n\\pagebreak\n")

# SMRs by age at diagnosis
cat(paste("\n Table", t + 1, " CVD SMRs by Age at Diagnosis for BC Survivors in Puerto Rico and US Race/Ethnicity Groups Using US population as Referent\n"))
t <- t + 1
gt_dx_agegroup

cat("\n\n\\pagebreak\n")

# SMRs by age at diagnosis figure
cat(paste("\n Figure", f + 1, " CVD SMRs by Age at Diagnosis for BC survivors in Puerto Rico and US Race/Ethnicity Groups Using US population as Referent\n"))
f <- f + 1
f_dx_agegroup
  
cat("\n\n\\pagebreak\n")

# SMRs by age at death
cat(paste("\n Table", t + 1, " CVD SMRs by Age at Death for BC Survivors in Puerto Rico and US Race/Ethnicity Groups Using US population as Referent\n"))
t <- t + 1
gt_agegroup

cat("\n\n\\pagebreak\n")

# SMRs by age at death figure
cat(paste("\n Figure", f + 1, " CVD SMRs by Age at Death for BC survivors in Puerto Rico and US Race/Ethnicity Groups Using US population as Referent\n"))
f <- f + 1
f_agegroup

cat("\n\n\\pagebreak\n")

# SMRs by stage
cat(paste("\n Table", t + 1, " CVD SMRs by Stage for BC Survivors in Puerto Rico and US Race/Ethnicity Groups Using US population as Referent\n"))
t <- t + 1
gt_stage

cat("\n\n\\pagebreak\n")

# SMRs by treatment
cat(paste("\n Table", t + 1, " CVD SMRs by Treatment for BC Survivors in Puerto Rico and US Race/Ethnicity Groups Using US population as Referent\n"))
t <- t + 1
gt_procedure

cat("\n\n\\pagebreak\n")

# SMRs by treatment and stage
cat(paste("\n Table", t + 1, " CVD SMRs by Stage and Treatment for BC Survivors in Puerto Rico and US Race/Ethnicity Groups Using US population as Referent\n"))
t <- t + 1
gt_proc_stage

cat("\n\n\\pagebreak\n")

# SMRs by treatment and stage figure
cat(paste("\n Figure", f + 1, " CVD SMRs by Stage and Treatment for BC survivors in Puerto Rico and US Race/Ethnicity Groups Using US population as Referent\n"))
f <- f + 1
f_proc_stage

cat("\n\n\\pagebreak\n")

# SMRs Table 1
cat(paste("\n Table", t + 1, " CVD SMRs for BC Survivors in Puerto Rico and US Race/Ethnicity Groups Using US population as Referent\n"))
t <- t + 1
gt_smr_table1

```

# Details

Figure legends should be an in-depth explanation of each figure, including a figure TITLE, and a CAPTION that includes the purpose of the figure, and brief method, results, and discussion statements pertaining to the figure. All abbreviations used in the figure should be identified either after their first mention in the legend or in alphabetical order at the end of each legend. All symbols used (arrows, circles, etc.) must be explained. Target length should be 50-100 words per figure.

*	All figures must have a number, title, and caption.

*	Figures should be cited in numerical order in the text.

*	Supplemental figures should be cited as “Online Figure 1, Online Figure 2,” etc.

*	Figure titles should be short and followed by a 2 to 3 sentence caption.

# Figures

*	Figures and graphs should be provided in EPS or TIF format.

*	Color images must be at least 300 DPI. Gray scale images should be at least 300 DPI.

*	All abbreviations used in the figure should be identified in an alphabetical order at the end of each legend.

*	All symbols used (arrows, circles, etc.) must be explained.

*	Figure legends should be typed double-spaced on pages separate from the text.

*	Figure numbers must correspond with the order in which they are mentioned in the text.

*	If previously published figures are used, written permission from the original publisher is required. See STM Guidelines for details: http://www.stm-assoc.org/copyright-legal-affairs/permissions/permissions-guidelines/.

*	If the figure has been previously published, cite the figure source in the legend.

Graphics software, such as Photoshop and Illustrator, should be used to create the art, but not presentation software such as PowerPoint, CorelDraw, or Harvard Graphics. Line art (black and white or color) and combinations of gray scale images and line art should be at least 1200 DPI. Lettering should be of sufficient size to be legible after reduction for publication. The optimal size is 12 points. Symbols should be of a similar size. Figures should be no smaller than 13 cm × 18 cm (500 × 700). Decimals, lines, and other details must be strong enough for reproduction. Use only black and white—not gray—in charts and graphs. Place crop marks on photomicrographs to show only the essential field. Designate special features with arrows. All symbols, arrows, and lettering on half-tone illustrations must contrast with the background. There is no fee for the publication of color figures. Our editors encourage authors to submit figures in color, as we feel it improves the clarity and visual impact of the images.


*	Your Central Illustration, if not an existing figure, should be listed first.

*	If the figure has been previously published, cite the figure source in the legend.

*	All abbreviations used in the figure should be identified in alphabetical order at the end of each legend (see also Figures).

# Tables

Each table should be on a separate page, with the table number and title centered above the table and explanatory notes below the table. Use Arabic numbers. Table numbers must correspond with the order cited in the text. Tables should be self-explanatory, and the data presented in them should not be duplicated in the text or figures.

*	All tables must have a title.

*	Abbreviations should be listed in a footnote under the table in alphabetical order.

*	Footnote symbols should appear in the following order: *,†,‡,§,‖,#,**,††, etc.

*	If previously published tables are used, written permission from the original publisher/author is required.

*	Cite the source of the table in the footnote.


```{r forest_plots, include = FALSE}

forest_format <- function(race_eth_select, ...) {
  # List all input dataframes
  dfs <- list(...)
  
  # Apply the transformation to each dataframe
  result <- dfs %>%
    lapply(function(df) {
      # Filter the dataframe by the specified race_eth value, rename columns, and apply formatting to Observed
      df %>%
        filter(race_eth == race_eth_select) %>%
        rename(Variable = names(df)[2]) %>%
        mutate(Variable = paste("    ", Variable)) %>%
        mutate(Observed = format(round(as.numeric(gsub(",", "", Observed)), 1), nsmall = 1)) %>%
        bind_rows(
          tibble(race_eth = NA, Variable = names(df)[2], Observed = NA, Expected = NA, SMR = NA, CI.Lower = NA, CI.Upper = NA)
        ) %>%
        arrange(desc(is.na(race_eth)), .by_group = TRUE)
    }) %>%
    bind_rows() %>%
    select(-race_eth) %>%
    mutate(` ` = paste(rep(" ", 20), collapse = " "),
           `SMR (95%)` = case_when(
             is.na(Observed) ~ "", 
             Observed == "-" ~ "", 
             TRUE ~ paste0(SMR, " (", CI.Lower, " - ", CI.Upper, ")")))
  
  return(result)
}

white_f <- forest_format(
  race_eth_select = "White", 
  merged_dx_agegroup, 
  merged_stage, 
  merged_procedure, 
  merged_g_deathyear, 
  merged_g_diagnosisyear
)

black_f <- forest_format(
  race_eth_select = "Black",  
  merged_dx_agegroup, 
  merged_stage, 
  merged_procedure, 
  merged_g_deathyear, 
  merged_g_diagnosisyear
)

us_hispanic_f <- forest_format(
  race_eth_select = "US Hispanic",  
  merged_dx_agegroup, 
  merged_stage, 
  merged_procedure, 
  merged_g_deathyear, 
  merged_g_diagnosisyear
)

pr_hispanic_f <- forest_format(
  race_eth_select = "PR Hispanic",  
  merged_dx_agegroup, 
  merged_stage, 
  merged_procedure, 
  merged_g_deathyear, 
  merged_g_diagnosisyear
)
aanhpi_f <- forest_format(
  race_eth_select = "AANHPI & AIAN", 
  merged_dx_agegroup, 
  merged_stage, 
  merged_procedure, 
  merged_g_deathyear, 
  merged_g_diagnosisyear
)

pr_hispanic_forest <- forest(pr_hispanic_f %>% select(-c(Observed, Expected, SMR, CI.Lower, CI.Upper)),
       est = as.numeric(pr_hispanic_f$SMR),
       lower = as.numeric(pr_hispanic_f$CI.Lower), 
       upper = as.numeric(pr_hispanic_f$CI.Upper),
       xlim = c(0, 25),
       ci_column = 2,
       ref_line = 1,
       xlab = "SMR",
       ticks_at = pretty(c(0, 25), n = 5),
       plot.margin=unit(c(-10,-10,-10,-10), "cm")); pr_hispanic_forest

us_hispanic_forest <- forest(us_hispanic_f %>% select(-c(Variable, Observed, Expected, SMR, CI.Lower, CI.Upper)),
       est = as.numeric(us_hispanic_f$SMR),
       lower = as.numeric(us_hispanic_f$CI.Lower), 
       upper = as.numeric(us_hispanic_f$CI.Upper),
       xlim = c(0, 25),
       ci_column = 1,
       ref_line = 1,
       xlab = "SMR",
       ticks_at = pretty(c(0, 25), n = 5),
       plot.margin=unit(c(-10,-10,-10,-10), "cm")); us_hispanic_forest

white_forest <- forest(white_f %>% select(-c(Variable, Observed, Expected, SMR, CI.Lower, CI.Upper)),
       est = as.numeric(white_f$SMR),
       lower = as.numeric(white_f$CI.Lower), 
       upper = as.numeric(white_f$CI.Upper),
       xlim = c(0, 25),
       ci_column = 1,
       ref_line = 1,
       xlab = "SMR",
       ticks_at = pretty(c(0, 25), n = 5),
       plot.margin=unit(c(-10,-10,-10,-10), "cm")); white_forest

black_forest <- forest(black_f %>% select(-c(Variable, Observed, Expected, SMR, CI.Lower, CI.Upper)),
       est = as.numeric(black_f$SMR),
       lower = as.numeric(black_f$CI.Lower), 
       upper = as.numeric(black_f$CI.Upper),
       xlim = c(0, 25),
       ci_column = 1,
       ref_line = 1,
       xlab = "SMR",
       ticks_at = pretty(c(0, 25), n = 5),
       plot.margin=unit(c(-10,-10,-10,-10), "cm")); black_forest

aanhpi_forest <- forest(aanhpi_f %>% select(-c(Variable, Observed, Expected, SMR, CI.Lower, CI.Upper)),
       est = as.numeric(aanhpi_f$SMR),
       lower = as.numeric(aanhpi_f$CI.Lower), 
       upper = as.numeric(aanhpi_f$CI.Upper),
       xlim = c(0, 25),
       ci_column = 1,
       ref_line = 1,
       xlab = "SMR",
       ticks_at = pretty(c(0, 25), n = 5),
       plot.background = element_rect(fill = "yellow"),
       plot.margin=unit(c(-10,-10,-10,-10), "cm")); aanhpi_forest

forest_figure <- grid.arrange(pr_hispanic_forest, us_hispanic_forest, white_forest, black_forest, aanhpi_forest, ncol = 5, clip = "on")

# knitr::plot_crop("/Users/genesisrodriguez/Downloads/Statistical Analysis/Outputs/Graphs/pr_hispanic_forest.tif")

ggsave("/Users/genesisrodriguez/Downloads/Statistical Analysis/Outputs/Graphs/forest_figure.tif", plot = forest_figure, width = 30, height = 8, dpi = 400,  bg = NULL)

beepr::beep(1)

```
