---
title: "Project: Cardiac death mortality on breast cancer patients (Dr. C Sanchez)"
author: "Edwin Navarro - BBC Biostatistician"
date: "2024-06-19"
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Working Directory -------------------------------------------------------

#setwd("C:/Users/enavarro/Documents/BBC/Project- Dra. Carola/Data")
#setwd("C:/Users/gabri/OneDrive/Documents/CCC/Project Dra. CSanchez/Data")
#setwd("C:/Users/lcastro/OneDrive - Centro Comprensivo de Cancer UPR/projects_shared/BBC/Project_Dra. Carola/Data")

# Libraries ---------------------------------------------------------------

{library(dplyr)
library(lubridate)
library(ggplot2)
library(tibble)
library(tidyr)
library(readxl)
library(stringr)
library(openxlsx)

library(survival)
library(survminer)
library(ggsurvfit)
library(gtsummary)
library(tidycmprsk)


library(knitr)
library(kableExtra)
library(RColorBrewer)}


```



```{r echo=FALSE, message=FALSE, warning=FALSE}

# Data --------------------------------------------------------------------

# Paths
paths <- c("C:/Users/Genesis Rodriguez/Box/Sanchez-Diaz Lab/CVD and BC/Registro Central de Cáncer/Data/datosCarola_invasive_seq0&1_NoID.xlsx",
           "~/Downloads/Registro Central de Cáncer/datosCarola_invasive_seq0&1_NoID.xlsx",
           paste0("C:/Users/lcastro/OneDrive - Centro Comprensivo de Cancer UPR/projects_shared/BBC/Project_Dra. Carola/Data/Registro Central de Cáncer/datosCarola_invasive_seq0&1_NoID.xlsx"))

# Loading dataframe

data <- suppressWarnings(tryCatch({
  read_excel(paths[1])
}, error = function(e) {
  read_excel(paths[2])
}, error = function(e) {
  read_excel(paths[3])
}))


# ER Assay, PR Assay & HER2 
# Reference links for encoding in data dictionary document
data <- data %>%
  mutate(ER = ifelse(ERassay  %in% c('1','010'), 'Positive', 
                     ifelse(ERassay %in% c('0','020'),'Negative',
                            ifelse(ERassay %in% c('999',NA,'9'), 'Unknown','Other'
                            ))),
         PR = ifelse(PRassay  %in% c('1','010'), 'Positive', 
                     ifelse(PRassay %in% c('0','020'),'Negative',
                            ifelse(PRassay %in% c('999',NA,'9'), 'Unknown','Other'
                            ))),
         HER2.cat = ifelse(HER2  %in% c('1','010'), 'Positive', 
                     ifelse(HER2 %in% c('0','020'),'Negative',
                            ifelse(HER2 %in% c('999',NA,'9'), 'Unknown','Other'
                            )))
         
  )
data$ER <- factor(data$ER, 
                       levels = c('Positive','Negative',
                                  'Other','Unknown'))
data$PR <- factor(data$PR, 
                       levels = c('Positive','Negative',
                                  'Other','Unknown'))
data$HER2.cat <- factor(data$HER2.cat, 
                       levels = c('Positive','Negative',
                                  'Other','Unknown'))

#Age Variable is character (set to numeric)
data$AgeDx <- as.numeric(data$AgeDx)

#Age Group Categories
data$AgeCat <- cut(data$AgeDx, breaks = c(0,17,59, 888,999), include.lowest = TRUE,
                   labels = c('0-17','18-59', '60+', 'Unknown'))

#Set factor order for Stage at Diagnosis
data$StageAtDx <- relevel(factor(data$StageAtDx), ref = "Localized")

# Health Region Variable
data$HealthRegion <- str_to_title(data$HealthRegion)
data$HealthRegion <- factor(data$HealthRegion,
                            levels = c('Noreste Metro','Central Bayamon','Oeste Aguadilla/Mayaguez',
                                       'Sureste Caguas','Sur Ponce',"Norte Arecibo",
                                       'Este Fajardo','County Unknown'
                                       ))

# Procedures (Treatment)

data <- data %>% 
  mutate(Procedure = ifelse(Surgery == 'Yes' & Chemotherapy != 'Yes' & Radiotherapy != 'Yes',
                         'Surgery Only', 
                 ifelse(Surgery != 'Yes' & Chemotherapy == 'Yes' & Radiotherapy != 'Yes',
                        'Chemotherapy Only',
                 ifelse(Surgery != 'Yes' & Chemotherapy != 'Yes' & Radiotherapy == 'Yes',
                        'Radiotheraphy Only',
                 ifelse(Surgery == 'Yes' & Chemotherapy == 'Yes' & Radiotherapy == 'Yes',
                        'Surgery + Chemotherapy + Radiotherapy',
                 ifelse(Surgery == 'Yes' & Chemotherapy == 'Yes' & Radiotherapy != 'Yes',
                        'Surgery + Chemotherapy',
                 ifelse(Surgery == 'Yes' & Chemotherapy != 'Yes' & Radiotherapy == 'Yes',
                        'Surgery + Radiotherapy',
                 ifelse(Surgery != 'Yes' & Chemotherapy == 'Yes' & Radiotherapy == 'Yes',
                        'Chemotherapy + Radiotherapy',
                 ifelse(Surgery == 'Unk' & Chemotherapy == 'Unk' & Radiotherapy == 'Unk',
                        'Unknown',
                 ifelse(Surgery == 'No' & Chemotherapy == 'No' & Radiotherapy == 'No',
                        'No treatment','No treatment'
                        )))))))))
)
data$Procedure <- with(data, factor(Procedure, 
                                    levels = c('Surgery + Chemotherapy + Radiotherapy',
                                               'Surgery + Radiotherapy','Surgery + Chemotherapy',
                                               'Chemotherapy + Radiotherapy',
                                               'Surgery Only','Chemotherapy Only', 
                                               'Radiotheraphy Only','No treatment','Unknown'))
)


# Comorbidities
data <- data %>%
  mutate(Comorb.cat = factor(ifelse(Comorb == '0' , 'No', 
                             ifelse(as.numeric(Comorb) >= 1, 'Yes', 'Unknown')),
                             levels= c('Yes','No','Unknown')) ,
         `Myocardial Infarction` = factor(ifelse(C1 == 0 & C0 == 0 , 'No', 
                             ifelse(C1 == 1 | C0 == 1, 'Yes', 'Unknown')),
                             levels= c('Yes','No','Unknown')), 
         `Congestive heart failure` = factor(ifelse(C2 == 0 , 'No', 
                             ifelse(C2 == 1, 'Yes', 'Unknown')),
                             levels= c('Yes','No','Unknown')),
         `Peripheral vascular disease` = factor(ifelse(C3 == 0 , 'No', 
                             ifelse(C3 == 1, 'Yes', 'Unknown')),
                             levels= c('Yes','No','Unknown')),
         `Cerebrovascular disease` = factor(ifelse(C4 == 0 , 'No', 
                             ifelse(C4 == 1, 'Yes', 'Unknown')),
                             levels= c('Yes','No','Unknown')),
         `Chronic pulmonary disease` = factor(ifelse(C6 == 0 , 'No', 
                             ifelse(C6 == 1, 'Yes', 'Unknown')),
                             levels= c('Yes','No','Unknown')),
         `Diabetes without chronic complication` = factor(ifelse(C10 == 0 , 'No', 
                             ifelse(C10 == 1, 'Yes', 'Unknown')),
                             levels= c('Yes','No','Unknown')),
         `Diabetes with chronic complication` = factor(ifelse(C11 == 0 , 'No', 
                             ifelse(C11 >= 1, 'Yes', 'Unknown')),
                             levels= c('Yes','No','Unknown')),
  )

# To determine and classify how many comorbidities 
data$Csum <- apply(data[, c('Myocardial Infarction','Congestive heart failure', 
               'Peripheral vascular disease',
               'Cerebrovascular disease',
               'Chronic pulmonary disease',
               'Diabetes without chronic complication',
               'Diabetes with chronic complication')],1, function(x)sum(grepl("Yes",x)))

data$CsumCat <- factor(with(data, ifelse(Csum == 0, 'No comorbidity',
                       ifelse(Csum >= 1 & Csum <=2, '1-2 comorbidities',
                       ifelse(Csum >= 3, '3 or more comorbidites', '')))),
                       levels = c('3 or more comorbidites','1-2 comorbidities','No comorbidity')
)



# CVD vs Non CVD death
data$Vital <- ifelse(data$VitalStatus == 0 & data$CVDDeathCause == 'CVD death','CVD death', 
                     ifelse(data$VitalStatus == 0 & data$BCDeathCause == 'BC death', 'BC death', 
                            ifelse(data$VitalStatus == 0, 'Other Death', 'Alive'
                            )))

# time: Delta year of diagnosis and year follow up (year of death if vital status == 0) 
# Vital Status: 1 = Alive, 0 = dead

# Recode status to 1 event (CVD death) and 0 censored
data$status <- ifelse(data$VitalStatus == 0 & data$CVDDeathCause == 'CVD death',1, 0)
data$status <- as.factor(data$status)

# Recode status to 1 event (CVD death) and 0 censored
data$statusBC <- ifelse(data$VitalStatus == 0 & data$BCDeathCause == 'BC death',1, 0)
data$statusBC <- as.factor(data$statusBC)

# Recode status to 1 event (All cause death) and 0 censored (for all cause mortality)
data$status1 <- ifelse(data$VitalStatus == 0 ,1, 0)

# Recode status to 1 (CVD death), 2 (Other death), 0 (Alive)
data <- data %>%
  mutate(
    status2 =  ifelse(data$VitalStatus == 0 & data$CVDDeathCause == 'CVD death',1, 
           ifelse(data$VitalStatus == 0, 2, 0
                  )),
    status2 = as.factor(status2),
    status2.cat = ifelse(status2 == 1, 'CVD Death',
                         ifelse(status2 == 2, 'Other Death', 'Alive'))
    )

# Recode status to 1 (CVD death), 2 (BC death), 3 (Other Death), 0 (Alive)
data <- data %>%
  mutate(
    statusV = ifelse(data$Vital == 'CVD death' ,1,
                     ifelse(data$Vital == 'BC death' ,2, 
           ifelse(data$Vital == 'Other Death', 3, 0
                  ))),
    statusV = as.factor(statusV),
    )

# time to follow up 
data <- data %>%
  mutate(time =yy_dlc-ydiag,
         AgeFU = AgeDx + time)


```

### **I. Research Objective:** 

Assess mortality on breast cancer survivors caused by cardiovascular disease (CVD) in women older than 17 years of age residing in Puerto Rico that are diagnosed with breast cancer (by microscopic confirmation) during the period of 2004-2019. 

#### <ins> A. Datasets (available at the moment):</ins> 

Cancer Registry Data: The initial selection is all women residents in PR diagnosed with breast cancer (by microscopic confirmation) during the period 2005-2019. 

#### <ins> B. Analysis Proposal: </ins>

Below is the overall proposed analysis strategy to meet the study objectives: 

1. Descriptive characteristics of the study population 

2. Estimate Standardized Mortality Ratios (SMR’s) with confidence intervals (Poisson-based CI’s) on the study population. Proposed SMR definition: observed number of CVD deaths among breast cancer patients divided by the expected number of CVD deaths in the matched general female population in PR.  

3. Estimate Cumulative Incidence of Death (Cumulative Mortality) considering the effects of competing risks (Cardiac-related death vs Other/Breast Cancer-related death).

### **II. Descriptive Analysis**

#### <ins> A. Summary Counts </ins>

*Table 1: Baseline Characteristics of Study Population* 
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Descriptive Analysis ----------------------------------------------------

# Alternative code
data %>%
  select(c(Vital, Procedure, StageAtDx, AgeCat, ER, PR, HER2.cat, CsumCat, `Myocardial Infarction`, `Congestive heart failure`, `Peripheral vascular disease`, `Cerebrovascular disease`, `Chronic pulmonary disease`, `Diabetes without chronic complication`, `Diabetes with chronic complication`, HealthRegion)) %>%
 tbl_summary(by = Vital,
              label = c(Procedure ~ "Treatment",
                        StageAtDx ~ "Stage at Diagnosis",
                        AgeCat ~ "Age at Diagnosis",
                        ER ~ "ER Assay",
                        PR ~ "PR Assay",
                        HER2.cat ~ "HER2", 
                        CsumCat ~ "Comorbidities",
                        HealthRegion ~ "Health Region"))  %>%
  add_overall() %>%
  modify_caption("**Table 1: Baseline Characteristics of Study Population**") %>%
  as_gt() %>%
  gt::tab_source_note("Note: Values are N (%), unless otherwise indicated.")

# Overall Count
# tab1 <- as.tibble(xtabs(~Procedure, data))
# tab1$Group <- 'Treatment'
# colnames(tab1) <- c('Variable', 'Freq', 'Group')
# tab2 <- as.tibble(xtabs(~StageAtDx, data))
# tab2$Group <- 'Stage at Diagnosis'
# colnames(tab2) <- c('Variable', 'Freq', 'Group')
# tab3 <- as.tibble(xtabs(~AgeCat, data))
# tab3$Group <- 'Age at Diagnosis'
# colnames(tab3) <- c('Variable', 'Freq', 'Group')
# tab4 <- as.tibble(xtabs(~ER, data))
# tab4$Group <- 'ER Assay'
# colnames(tab4) <- c('Variable', 'Freq', 'Group')
# tab5 <- as.tibble(xtabs(~PR, data))
# tab5$Group <- 'PR Assay'
# colnames(tab5) <- c('Variable', 'Freq', 'Group')
# tab6 <- as.tibble(xtabs(~HER2.cat, data))
# tab6$Group <- 'HER2'
# colnames(tab6) <- c('Variable', 'Freq', 'Group')
# tab8 <- as.tibble(xtabs(~CsumCat, data))
# tab8$Group <- 'Comorbidities'
# colnames(tab8) <- c('Variable', 'Freq', 'Group')
# tab9 <- as.tibble(xtabs(~`Myocardial Infarction`, data))
# tab9$Group <- 'Myocardial Infarction'
# colnames(tab9) <- c('Variable', 'Freq', 'Group')
# tab10 <- as.tibble(xtabs(~`Congestive heart failure`, data))
# tab10$Group <- 'Congestive heart failure'
# colnames(tab10) <- c('Variable', 'Freq', 'Group')
# tab11 <- as.tibble(xtabs(~`Peripheral vascular disease`, data))
# tab11$Group <- 'Peripheral vascular disease'
# colnames(tab11) <- c('Variable', 'Freq', 'Group')
# tab12 <- as.tibble(xtabs(~`Cerebrovascular disease`, data))
# tab12$Group <- 'Cerebrovascular disease'
# colnames(tab12) <- c('Variable', 'Freq', 'Group')
# tab13 <- as.tibble(xtabs(~`Chronic pulmonary disease`, data))
# tab13$Group <- 'Chronic pulmonary disease'
# colnames(tab13) <- c('Variable', 'Freq', 'Group')
# tab14 <- as.tibble(xtabs(~`Diabetes without chronic complication`, data))
# tab14$Group <- 'Diabetes without chronic complication'
# colnames(tab14) <- c('Variable', 'Freq', 'Group')
# tab15 <- as.tibble(xtabs(~`Diabetes with chronic complication`, data))
# tab15$Group <- 'Diabetes with chronic complication'
# colnames(tab15) <- c('Variable', 'Freq', 'Group')
# tab16 <- as.tibble(xtabs(~HealthRegion, data))
# tab16$Group <- 'Health Region'
# colnames(tab16) <- c('Variable','Freq', 'Group')
# 
# tab1 <- tab1%>%
#   bind_rows(tab2,tab3,tab4,tab5,tab6,tab8,tab9,tab10,tab11,tab12,tab13,tab14,tab15,
#             tab16)
#   
# rm(tab2);rm(tab3);rm(tab4);rm(tab5);rm(tab6);
# rm(tab8);rm(tab9);rm(tab10);rm(tab11);rm(tab12)
# rm(tab13);rm(tab14);rm(tab15) 
# 
# tab1 <- tab1 %>%
#   group_by(Group) %>% 
#   mutate(Percent = round((Freq/ sum(Freq))*100,2),
#          Value = paste0(Freq, '(', Percent, '%', ')')
#          )%>%
#   select(Group, Variable, Value)
# 
# 
# # Count by Vital Status
#   
# tab1.1 <- as.tibble(xtabs(~Procedure+Vital, data))
# tab1.1$Group <- 'Treatment'
# colnames(tab1.1) <- c('Variable','Vital','Freq', 'Group')
# tab2.2 <- as.tibble(xtabs(~StageAtDx+Vital, data))
# tab2.2$Group <- 'Stage at Diagnosis'
# colnames(tab2.2) <- c('Variable','Vital','Freq', 'Group')
# tab3.3 <- as.tibble(xtabs(~AgeCat+Vital, data))
# tab3.3$Group <- 'Age at Diagnosis'
# colnames(tab3.3) <- c('Variable','Vital','Freq', 'Group')
# tab4.4 <- as.tibble(xtabs(~ER+Vital, data))
# tab4.4$Group <- 'ER Assay'
# colnames(tab4.4) <- c('Variable','Vital','Freq', 'Group')
# tab5.5 <- as.tibble(xtabs(~PR+Vital, data))
# tab5.5$Group <- 'PR Assay'
# colnames(tab5.5) <- c('Variable','Vital','Freq', 'Group')
# tab6.6 <- as.tibble(xtabs(~HER2.cat+Vital, data))
# tab6.6$Group <- 'HER2'
# colnames(tab6.6) <- c('Variable','Vital','Freq', 'Group')
# tab8.8 <- as.tibble(xtabs(~CsumCat+Vital, data))
# tab8.8$Group <- 'Comorbidities'
# colnames(tab8.8) <- c('Variable','Vital', 'Freq', 'Group')
# tab9.9 <- as.tibble(xtabs(~`Myocardial Infarction`+Vital, data))
# tab9.9$Group <- 'Myocardial Infarction'
# colnames(tab9.9) <- c('Variable', 'Vital','Freq', 'Group')
# tab10.1 <- as.tibble(xtabs(~`Congestive heart failure`+Vital, data))
# tab10.1$Group <- 'Congestive heart failure'
# colnames(tab10.1) <- c('Variable', 'Vital','Freq', 'Group')
# tab11.1 <- as.tibble(xtabs(~`Peripheral vascular disease`+Vital, data))
# tab11.1$Group <- 'Peripheral vascular disease'
# colnames(tab11.1) <- c('Variable','Vital', 'Freq', 'Group')
# tab12.1 <- as.tibble(xtabs(~`Cerebrovascular disease`+Vital, data))
# tab12.1$Group <- 'Cerebrovascular disease'
# colnames(tab12.1) <- c('Variable', 'Vital','Freq', 'Group')
# tab13.1 <- as.tibble(xtabs(~`Chronic pulmonary disease`+Vital, data))
# tab13.1$Group <- 'Chronic pulmonary disease'
# colnames(tab13.1) <- c('Variable','Vital', 'Freq', 'Group')
# tab14.1 <- as.tibble(xtabs(~`Diabetes without chronic complication`+Vital, data))
# tab14.1$Group <- 'Diabetes without chronic complication'
# colnames(tab14.1) <- c('Variable', 'Vital','Freq', 'Group')
# tab15.1 <- as.tibble(xtabs(~`Diabetes with chronic complication`+Vital, data))
# tab15.1$Group <- 'Diabetes with chronic complication'
# colnames(tab15.1) <- c('Variable', 'Vital','Freq', 'Group')
# tab16.1 <- as.tibble(xtabs(~HealthRegion+Vital, data))
# tab16.1$Group <- 'Health Region'
# colnames(tab16.1) <- c('Variable', 'Vital','Freq', 'Group')
# 
# 
# tab1.1 <- tab1.1%>%
#   bind_rows(tab2.2,tab3.3,tab4.4,tab5.5,tab6.6,tab8.8,tab9.9,tab10.1,tab11.1,
#             tab12.1,tab13.1,tab14.1,tab15.1,tab16.1)
#   
# rm(tab2.2);rm(tab3.3);rm(tab4.4);rm(tab5.5);rm(tab6.6)  
# rm(tab8.8);rm(tab9.9);rm(tab10.1);rm(tab11.1);rm(tab12.1)
# rm(tab13.1);rm(tab14.1);rm(tab15.1) 
# 
# 
# tab1.1 <- tab1.1 %>%
#   group_by(Group, Vital) %>% 
#   mutate(Percent = round((Freq/ sum(Freq))*100,2),
#          Value = paste0(Freq, '(', Percent, '%', ')')
#          )%>% 
#   select(-c(Freq, Percent))%>%
#   pivot_wider(names_from = Vital, values_from = Value, values_fill = '0(0%)')%>%
#   select(c(Group, Variable,Alive, `CVD death`, `BC death`, `Other Death`))
# 
# 
# # Final Table
# 
# tabF <- tab1 %>% inner_join(tab1.1, by = c('Variable','Group'))
# rm(tab1.1);rm(tab1)
# 
# tabF[,2:dim(tabF)[2]]%>%
#   kable(format = "html", escape = FALSE, align = "c", col.names = c("", "Total (N = 27238)",
#                                            "Alive (N=18943)","Cardiac Death (N=537)","Breast Cancer Death (N=4595)" , "Other Cause Death (N=3163)"))%>%
#   group_rows(index = setNames(rle(tabF$Group)[[1]], rle(tabF$Group)[[2]]))%>%
#   kable_styling( full_width = FALSE) %>%
#   column_spec(1, bold = TRUE) %>%
#   footnote(general = "Values are N (%), unless otherwise indicated.",
#            general_title = "Note: ", 
#            footnote_as_chunk = TRUE)
```

*Table 2: Summary statistics by mortality type*

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Alternative code
data %>%
  filter(Vital != "Alive", AgeDx != 999, time >=0 ) %>% 
  select(c(Vital, AgeDx, AgeFU, time)) %>%
 tbl_summary(by = Vital,
              label = c(AgeDx ~ "Age at Diagnosis",
                       AgeFU ~ "Age at Death",
                       time ~ "Time to death (years)"),
             type = all_continuous() ~ "continuous2",
             statistic = list(all_continuous() ~ c(
      "{mean}(±{sd})",
      "{median} ({p25}-{p75})",
      "{min}-{max}"),
    all_categorical() ~ "{n} ({p}%)"))  %>%
  modify_caption("**Table 2: Summary statistics by mortality type**")



# tab2 <- data %>%
#   filter(Vital != "Alive", AgeDx != 999, time >=0 )%>% 
#   group_by(Vital)%>%
#   summarize(`Average Age at Diagnosis`= paste0(round(mean(AgeDx),2),"(","\u00B1",round(sd(AgeDx),2),")"),
#             `Median Age at Diagnosis`= paste0(round(median(AgeDx),2),"(",
#                                                      quantile(AgeDx,.25),"-",quantile(AgeDx,.75),")"),
#             `Range Age at Diagnosis` = paste0(range(AgeDx)[1],'-',range(AgeDx)[2]),
#             `Average Age at Death`= paste0(round(mean(AgeFU),2),"(","\u00B1",round(sd(AgeFU),2),")"),
#             `Median Age at Death`= paste0(round(median(AgeFU),2),"(",
#                                                      quantile(AgeFU,.25),"-",quantile(AgeFU,.75),")"),
#             `Range Age at Death` = paste0(range(AgeFU)[1],'-',range(AgeFU)[2]),
#             `Average Time to death (years)`  = paste0(round(mean(time),2),"(","\u00B1",round(sd(time),2),")"),
#             `Median Time to death (years)`  = paste0(round(median(time),2),"(",
#                                                      quantile(time,.25),"-",quantile(time,.75),")"),
#             `Range Time to death (years)` = paste0(range(time)[1],'-',range(time)[2])
#   )%>%  
#   pivot_longer(cols = !Vital, names_to = "Variable", values_to = "Metric")%>%
#   pivot_wider(names_from = Vital, values_from = Metric)%>%
#   mutate(group = ifelse(str_detect(Variable, 'Diagnosis'), 'Age at Diagnosis',
#                  ifelse(str_detect(Variable, 'at Death'), 'Age at Death','Time to death (years)'
#                         )),
#          Variable = str_split_i(Variable," ",1)
#          )%>%
#   select(group, Variable, `BC death`, `CVD death`,`Other Death`)
# 

# tab2[,2:dim(tab2)[2]] %>%
#   kable(format = "html", escape = FALSE, align = "c", col.names = c("", "Breast Cancer Death (N=4595)",
#                                            "Cardiac Death (N=537)", "Other Cause Death (N=3163)"))%>%
#   group_rows(index = setNames(rle(tab2$group)[[1]], rle(tab2$group)[[2]]))%>%
#   kable_styling( full_width = FALSE) %>%
#   column_spec(1, bold = TRUE)
```



#### <ins> B. Exploratory Data Analysis </ins>  

*Figure: Age distribution of study population*

```{r echo=FALSE, message=FALSE, warning=FALSE}

data %>%
  filter(AgeDx != 999)%>%
  ggplot(aes(x = AgeDx))+
  geom_histogram()+
  labs(x = 'Age at diagnosis', y = 'Frequency',
       title = 'Age distribution')+
  theme_bw()
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
data %>%
  filter(AgeDx != 999)%>%
  ggplot(aes(x = AgeDx))+
  geom_histogram()+
  #facet_wrap(~StageAtDx,scales = 'free')+
  facet_wrap(~Vital, scales = 'free_y')+
  labs(x = 'Age at diagnosis', y = 'Frequency',
       title = 'Age distribution by mortality type')+
  theme_bw()
```
  
  
*Figure: Bar plots for procedures (surgery,chemotherapy,radiotherapy)  *

```{r echo=FALSE, message=FALSE, warning=FALSE}

par(mfrow=c(1,3))
barplot(sort(xtabs(~Surgery, data = data), decreasing = TRUE),  main = 'Surgery')
barplot(sort(xtabs(~Chemotherapy, data = data), decreasing = TRUE), main = 'Chemotherapy')
barplot(sort(xtabs(~Radiotherapy, data = data), decreasing = TRUE), main = 'Radiotherapy')

# data %>% 
#  group_by(Vital, Surgery) %>%
#  summarize(Frequency = n())%>%
#  ggplot(aes(x = reorder(Surgery,-Frequency), y = Frequency))+
#    geom_bar(stat = 'identity')+
#    labs(x = 'Surgery')+
#    geom_text(aes(label = Frequency), vjust = 0) +
#    facet_wrap(~Vital, scales = "free")+
#    theme_minimal()

``` 


```{r echo=FALSE, fig.height=10, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
# plotting a barplot with percentages 

data %>%  
  group_by(Vital, Procedure) %>%
  summarize(Frequency = n())%>% 
  mutate(Percent = prop.table(Frequency)) %>% 
  ungroup() %>% 
  ggplot(aes(x = reorder(Procedure,-Percent), y = Percent, label = scales::percent(Percent))) +  
  geom_bar(stat = 'identity',color = 'black') +  
  geom_text(vjust = -0.5,    # nudge above top of bar
            size = 3) +  
  facet_wrap(~Vital)+
  theme_bw()+
  labs(x = '')+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_y_continuous(labels = scales::percent)

```

```{r echo=FALSE, fig.height=6, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
data %>%  
  group_by(Vital, Procedure) %>%
  summarize(Frequency = n())%>% 
  mutate(Percent = round(prop.table(Frequency),3)) %>% 
  ungroup() %>% 
  ggplot(aes(x = reorder(Procedure,-Percent), y = Percent, fill = Vital)) + 
  #ggplot(aes(x = Procedure, y = Percent, fill = Vital)) +  
  geom_bar(stat = 'identity',color = 'black', position = 'dodge') +  
  labs(x = '')+
  #facet_wrap(~Procedure, scales = "free_x")+
  scale_fill_grey(start=0, end=1)+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.title = element_blank())+
  scale_y_continuous(labels = scales::percent)

```

  
*Figure: Bar plot for Comorbidity Index  *

```{r echo=FALSE, message=FALSE, warning=FALSE}
barplot(xtabs(~data$Comorb))
```

  
*Figure: Bar plots for ER, PR & HER2  *

```{r echo=FALSE, fig.height=6, fig.width=10, message=FALSE, warning=FALSE}

par(mfrow=c(1,3))
barplot(xtabs(~ER, data = data), main = 'ER Assay')
barplot(xtabs(~PR, data = data), main = 'PR Assay')
barplot(xtabs(~HER2.cat, data = data), main = 'HER2')

```


```{r echo=FALSE, fig.height=6, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}

## ER
data %>%  
  group_by(Vital, ER) %>%
  summarize(Frequency = n())%>% 
  mutate(Percent = round(prop.table(Frequency),3)) %>% 
  ungroup() %>% 
  ggplot(aes(x = reorder(ER,-Percent), y = Percent, fill = Vital)) + 
  #ggplot(aes(x = ER, y = Percent, fill = Vital)) +  
  geom_bar(stat = 'identity',color = 'black', position = 'dodge') +  
  labs(x = '', title = "ER")+
  scale_fill_grey(start=0, end=1)+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.title = element_blank())+
  scale_y_continuous(labels = scales::percent)

## PR
data %>%  
  group_by(Vital, PR) %>%
  summarize(Frequency = n())%>% 
  mutate(Percent = round(prop.table(Frequency),3)) %>% 
  ungroup() %>% 
  ggplot(aes(x = reorder(PR,-Percent), y = Percent, fill = Vital)) + 
  #ggplot(aes(x = ER, y = Percent, fill = Vital)) +  
  geom_bar(stat = 'identity',color = 'black', position = 'dodge') +  
  labs(x = '', title = "PR")+
  scale_fill_grey(start=0, end=1)+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.title = element_blank())+
  scale_y_continuous(labels = scales::percent)

## HER2
data %>%  
  group_by(Vital, HER2.cat) %>%
  summarize(Frequency = n())%>% 
  mutate(Percent = round(prop.table(Frequency),3)) %>% 
  ungroup() %>% 
  ggplot(aes(x = reorder(HER2.cat,-Percent), y = Percent, fill = Vital)) + 
  #ggplot(aes(x = ER, y = Percent, fill = Vital)) +  
  geom_bar(stat = 'identity',color = 'black', position = 'dodge') +  
  labs(x = '', title = 'HER2')+
  scale_fill_grey(start=0, end=1)+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.title = element_blank())+
  scale_y_continuous(labels = scales::percent)
```

### **III. All Cause Mortality**

*Figure 1: All Cause Mortality Survival Curve* 

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Fit the Kaplan-Meier survival model

allC <- data %>%
  select(time,status1)

km_fit <- survfit(Surv(time, status1)~1, allC)

km_plot <- ggsurvplot(
  km_fit, 
  data = data,
  risk.table = TRUE,       # Add number at risk table
 # pval = TRUE,             # Show p-value of log-rank test (although not relevant here)
  conf.int = TRUE,         # Show confidence intervals
  xlab = "Years",          # X-axis label
  ylab = "Probability of Survival", # Y-axis label
  ggtheme = theme_minimal() # Theme
)

km_plot$plot <- km_plot$plot +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  ) +
  labs(title = "Overall Survival Curve")

# Print the plot
print(km_plot)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
allC <- tidy(km_fit)

allC  <- allC   %>%
  filter(time %in% c(5,10,15)) %>%
  mutate(value= paste0(100 * round(estimate,4), '(',100*round(conf.low,4), '-',100*round(conf.high,4), ')' )
        # outcome = ifelse(outcome == '1','CVD Death',
        #                  ifelse(outcome == '2', 'Other Cause Death',''))
        )%>%
  select(time, value)%>%
  pivot_wider(values_from = value, names_from = time)%>%
  rename(`5-year (95% CI)` = `5`,
         `10-year (95% CI)` = `10`,
         `15-year (95% CI)` = `15`)
  
kable(allC,  format = "html", escape = FALSE, align = "c")%>%
  kable_styling( full_width = FALSE) %>%
  column_spec(1, bold = TRUE) %>%
  footnote(general = "Values are n (%), unless otherwise indicated. Crude number and rate estimates of mortality from the study data. All-cause mortality estimated by Kaplan-Meier methods",
           general_title = "", 
           footnote_as_chunk = TRUE)

```

*Figure 2: All Cause Mortality Survival Curve by Treatment* 

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Fit the Kaplan-Meier survival model

km_fit2 <- survfit(Surv(time, status1)~Procedure, data = data)

km_plot2 <- ggsurvplot(
  km_fit2, 
  data = data,
  risk.table = FALSE,       # Add number at risk table
  pval = TRUE,             # Show p-value of log-rank test (although not relevant here)
 # conf.int = TRUE,         # Show confidence intervals
  xlab = "Years",          # X-axis label
  ylab = "Probability of Survival", # Y-axis label
   legend.title = "Procedure",  # Legend title
 
    legend.labs = levels(data$Procedure),
  palette = c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#C695DD","#E25869","#95CD9E","#8D3CE5"), # Custom color palette
  ggtheme = theme_minimal() 
  ) 

km_plot2$plot <- km_plot2$plot +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8)
  ) +
  labs(title = "Survival Curve by Treatment") +
  guides(color = guide_legend(ncol = 3)) 

# Print the plot
print(km_plot2)

```

*Figure 2a: All Cause Mortality Survival Curve by Surgery Treatment* 

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Fit the Kaplan-Meier survival model
surgery_data <- data %>%
  filter(grepl("Surgery", Procedure))
surgery_data$Procedure <- factor(surgery_data$Procedure)

km_fit2 <- survfit(Surv(time, status1)~Procedure, data = surgery_data)

km_plot2 <- ggsurvplot(
  km_fit2, 
  data = surgery_data,
  risk.table = FALSE,       # Add number at risk table
  pval = TRUE,             # Show p-value of log-rank test (although not relevant here)
 # conf.int = TRUE,         # Show confidence intervals
  xlab = "Years",          # X-axis label
  ylab = "Probability of Survival", # Y-axis label
   legend.title = "Procedure",  # Legend title
    legend.labs = levels(surgery_data$Procedure),
  palette = c("#E69F00", "#56B4E9", "#009E73",  "#0072B2"), # Custom color palette
  ggtheme = theme_minimal() 
  ) 

km_plot2$plot <- km_plot2$plot +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8)
  ) +
  labs(title = "Survival Curve by Survery Treatment") +
  guides(color = guide_legend(ncol = 3)) 

# Print the plot
print(km_plot2)

```


```{r echo=FALSE, message=FALSE, warning=FALSE}

allCT <- tidy(km_fit2)
allCT[is.na(allCT)] <- 0

allCT  <- allCT   %>%
  filter(time %in% c(5,10,15)) %>%
  mutate(value= paste0(100 * round(estimate,4), '(',100*round(conf.low,4), '-',100*round(conf.high,4), ')' ),
        strata = gsub('Procedure=','',strata))%>%
  select(time, strata,value)%>%
  pivot_wider(values_from = value, names_from = time, values_fill = '0')%>%
  rename(`5-year (95% CI)` = `5`,
         `10-year (95% CI)` = `10`,
         `15-year (95% CI)` = `15`,
         Treatment = strata)
  
kable(allCT,  format = "html", escape = FALSE, align = "c")%>%
  kable_styling( full_width = FALSE) %>%
  column_spec(1, bold = TRUE) %>%
  footnote(general = "Values are n (%), unless otherwise indicated. Crude number and rate estimates of mortality from the study data. All-cause mortality estimated by Kaplan-Meier methods",
           general_title = "", 
           footnote_as_chunk = TRUE)

rm(allC);rm(allCT);rm(km_fit);rm(km_fit2);rm(km_plot);rm(km_plot2)
```

### **IV. Competing Risks**

If the goal is to assess cardiac death mortality, both death related to breast cancer 
and other causes are competing events.   

*Figure 1.A: Cumulative incidence functions (1 = Cardiac death ; 2 = Other cause death)* 
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Calculating CIF ----------------------------------------------

# Estimate the cumulative incidence in the context of competing risks using the 
# cuminc function from the {tidycmprsk} package. By default this requires the
# status to be a factor variable with censored patients coded as 0.

#cuminc(Surv(time, status2) ~ 1, data = data)

# We can use the ggcuminc() function from the {ggsurvfit} package to plot the 
# cumulative incidence. By default it plots the first event type only. 
# So the following plot shows the cumulative incidence of death from cvd:

# include both event outcomes
cuminc(Surv(time, status2)~1, data) %>% 
  ggcuminc(outcome = c(  "1", "2"))+
  ylim(c(0, 1)) + 
  labs(
    x = "Survival Time (Years)"
  ) + 
  xlim(c(0,20))+
  add_confidence_interval()



```
  
  
*Table: Cumulative Incidence of Death (CVD Death vs Other)*

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

way2 <- tidy(cuminc(Surv(time, status2)~1, data))

way2 <- way2 %>%
  filter(time %in% c(5,10,15)) %>%
  mutate(value= paste0(100 * round(estimate,4), '(',100*round(conf.low,4), '-',100*round(conf.high,4), ')' ),
         outcome = ifelse(outcome == '1','CVD Death',
                          ifelse(outcome == '2', 'Other Cause Death','')))%>%
  select(time, outcome, value)%>%
  pivot_wider(values_from = value, names_from = time)%>%
  rename(`5-year (95% CI)` = `5`,
         `10-year (95% CI)` = `10`,
         `15-year (95% CI)` = `15`,
         ` ` = outcome)
  
kable(way2,  format = "html", escape = FALSE, align = "c")%>%
  kable_styling( full_width = FALSE) %>%
  column_spec(1, bold = TRUE) %>%
  footnote(general = "Values are n (%), unless otherwise indicated. CVD and BC mortality are estimated by Fine and Gray’s competing risk method.",
           general_title = "Note: ", 
           footnote_as_chunk = TRUE)

```
  
  
*Figure 1.B: Cumulative incidence functions (1 = Cardiac death ; 2 = Breast Cancer death, 3 = Other Cause death)* 

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Calculating CIF ----------------------------------------------


# include both event outcomes

cuminc(Surv(time, statusV)~1, data) %>% 
  ggcuminc(outcome = c( "1", "2","3"),
           linetype_aes = TRUE)+
  ylim(c(0, .5)) + 
  labs(
    x = "Survival Time (Years)"
  ) + 
  xlim(c(0,20))+
  scale_linetype_manual(values=c( "solid", "longdash", "dotted"))


```

  
  
*Table: Cumulative Incidence of Death (including BC mortality)*
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

way3 <- tidy(cuminc(Surv(time, statusV)~1, data))

way3 <- way3 %>%
  filter(time %in% c(5,10,15)) %>%
  mutate(value= paste0(100 * round(estimate,4), '(',100*round(conf.low,4), '-',100*round(conf.high,4), ')' ),
         outcome = ifelse(outcome == '1','CVD Death',
                          ifelse(outcome == '2', 'BC Death', 'Other Cause Death')))%>%
  select(time, outcome, value)%>%
  pivot_wider(values_from = value, names_from = time)%>%
  rename(`5-year (95% CI)` = `5`,
         `10-year (95% CI)` = `10`,
         `15-year (95% CI)` = `15`,
         ` ` = outcome)
  
kable(way3,  format = "html", escape = FALSE, align = "c")%>%
  kable_styling( full_width = FALSE) %>%
  column_spec(1, bold = TRUE) %>%
  footnote(general = "Values are n (%), unless otherwise indicated. CVD, BC mortality and Other Cause mortality are estimated by Fine and Gray’s competing risk method.",
           general_title = "Note: ", 
           footnote_as_chunk = TRUE)

```

#### **IV. Cumulative Incidence Curves for CVD Mortality **

*Figure 1.B: Cumulative incidence curves for CVD Mortality* 


```{r echo=FALSE, fig.height=8, fig.width=12, message=FALSE, warning=FALSE, paged.print=FALSE}

# CIF for CVD by Treatment 

cuminc(Surv(time, status2) ~ Procedure, data) %>% 
  ggcuminc(outcome = c( "1"),
           linetype_aes = TRUE, size = 1)+
 #add_confidence_interval() +
 #add_risktable() +
  ylim(c(0, .1)) + 
  labs(
    x = "Survival Time (Years)",
    title = 'CIF for CVD by Treatment'
  ) + 
  xlim(c(0,20)) +  scale_linetype_manual(values=c( 1:9))


```


```{r echo=FALSE, fig.height=8, fig.width=12, message=FALSE, warning=FALSE, paged.print=FALSE}

# CIF for CVD by Age Category 

cuminc(Surv(time, status2) ~ AgeCat, data) %>% 
  ggcuminc(outcome = c( "1"),
           linetype_aes = TRUE, size = 1)+
 #add_confidence_interval() +
 #add_risktable() +
  ylim(c(0, .1)) + 
  labs(
    x = "Survival Time (Years)",
    title = 'CIF for CVD by Age at Diagnosis'
  ) + 
  xlim(c(0,20)) +  scale_linetype_manual(values=c( 1:9))


```


```{r echo=FALSE,fig.height=8, fig.width=12, message=FALSE, warning=FALSE, paged.print=FALSE}

# CIF for CVD by Comorbidity 

cuminc(Surv(time, statusV) ~ Comorb.cat , data) %>% 
  ggcuminc(outcome = c("1"),
           linetype_aes = TRUE, size = 1)+
 #add_confidence_interval() +
 #add_risktable() +
  ylim(c(0, .1)) + 
  labs(
    x = "Survival Time (Years)",
    title = 'CIF for Cardiovascular Disease by Comorbidity'
  ) + 
  xlim(c(0,20)) + scale_linetype_manual(values=c(rep("dashed",3)))


# CIF for BC by Comorbidity 

cuminc(Surv(time, statusV) ~ Comorb.cat , data) %>% 
  ggcuminc(outcome = c("2"),
           linetype_aes = TRUE, size = 1)+
 #add_confidence_interval() +
 #add_risktable() +
  labs(
    x = "Survival Time (Years)",
    title = 'CIF for Breast Cancer by Comorbidity'
  ) + 
  xlim(c(0,20)) + scale_linetype_manual(values=c(rep("dotted",3)))


# CIF for CVD by Comorbidity 

cuminc(Surv(time, statusV) ~ Comorb.cat , data) %>% 
  ggcuminc(outcome = c("1", "2"),
           linetype_aes = TRUE, size = 1)+
 #add_confidence_interval() +
 #add_risktable() +
  #ylim(c(0, .1)) + 
  labs(
    x = "Survival Time (Years)",
      title = 'CIF for Cardiovascular Disease and Breast Cancer by Comorbidity'
  ) + 
  xlim(c(0,20)) +   scale_linetype_manual(values = c("dashed", "dotted"),
                        labels = c("Cardiovascular disease", "Breast Cancer"))


```


```{r echo=FALSE, fig.height=8, fig.width=12, message=FALSE, warning=FALSE, paged.print=FALSE}

# CIF for CVD by cancer Stage at Diagnosis 

cuminc(Surv(time, status2) ~ StageAtDx , data) %>% 
  ggcuminc(outcome = c( "1"),
           linetype_aes = TRUE, size = 1)+
 #add_confidence_interval() +
 #add_risktable() +
  ylim(c(0, .1)) + 
  labs(
    x = "Survival Time (Years)",
    title = 'CIF for CVD by Stage at Diagnosis'
  ) + 
  xlim(c(0,20)) +  scale_linetype_manual(values=c( 1:9))


```

#### <ins>A. Competing Risks Regression </ins> 

##### 1. Subdistribution of Harzards 


*Table : Hazard ratios (and 95% Confidence Intervals) from Subdistribution Hazard Model for Cardiac Death* 
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Competing Risks Regression ----------------------------------------------



model <- crr(Surv(time,status2) ~ StageAtDx +Comorb.cat+
                         Procedure, data = data)

model %>%
   tbl_regression(exp = TRUE)


```



#### Forest Plot of Hazard Ratios

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Extract the results
model_summary <- summary(model)
tidy_model <- tidy(model)

# Create a data frame of the results
hr_results <- data.frame(
  group = tidy_model$term ,
  HR = exp(tidy_model$estimate ),
  lower_CI = exp(tidy_model$estimate - 1.96 * tidy_model$std.error),
  upper_CI = exp(tidy_model$estimate + 1.96 * tidy_model$std.error)
)


## Forest Plot for Stage at Diagnosis
# Filter the results for the group variable
hr_stage <- hr_results %>% filter(grepl("StageAtDx", group))

# Rename the group variable for better readability
hr_stage$group <- gsub("StageAtDx", "", hr_stage$group)


# Create a forest plot
ggplot(hr_stage, aes(x = HR, y = group)) +
  geom_point() +
  geom_errorbarh(aes(xmin = lower_CI, xmax = upper_CI), height = 0.2) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(title = "Subdistribution HR (95% CI) by Stage at Diagnosis", x = "Hazard Ratio", y = "") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 12),
    plot.title = element_text(size = 14, face = "bold")
  )


## Forest Plot for Procedure

# Filter the results for the group variable
hr_trt <- hr_results %>% filter(grepl("Procedure", group))

# Rename the group variable for better readability
hr_trt$group <- gsub("Procedure", "", hr_trt$group)


# Create a forest plot
ggplot(hr_trt, aes(x = HR, y = group)) +
  geom_point() +
  geom_errorbarh(aes(xmin = lower_CI, xmax = upper_CI), height = 0.2) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(title = "Subdistribution HR (95% CI) by Treatment", x = "Hazard Ratio", y = "") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 12),
    plot.title = element_text(size = 14, face = "bold")
  )

## Forest Plot for Comorbidity

# Filter the results for the group variable
hr_cmb <- hr_results %>% filter(grepl("Comorb.cat", group))

# Rename the group variable for better readability
hr_cmb$group <- gsub("Comorb.cat", "", hr_cmb$group)


# Create a forest plot
ggplot(hr_cmb, aes(x = HR, y = group)) +
  geom_point() +
  geom_errorbarh(aes(xmin = lower_CI, xmax = upper_CI), height = 0.2) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(title = "Subdistribution HR (95% CI) by Comorbidity", x = "Hazard Ratio", y = "") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 12),
    plot.title = element_text(size = 14, face = "bold")
  )


```